---
title: "Formatting messages with cli  :: Cheatsheet"
description: " "
image-alt: ""
execute:
  eval: true
  output: true
  warning: true
  error: true
---

```{r setup, include = FALSE}
options(cli.num_colors = 256)
fansi::set_knit_hooks(knitr::knit_hooks)
```

<!-- overall structure: https://github.com/r-lib/cli/issues/544 -->

<!--- rendering issue: https://github.com/quarto-dev/quarto-cli/discussions/3957 -->

cli offers a range of tools for building attractive command line interfaces (CLIs). This cheatsheet covers features of cli useful for raising errors, warnings and messages in R functions, and printing messages with rich text formatting to the console. For more advanced features, see the [cli package docs](https://cli.r-lib.org)

```{r}
library(cli)
```

## Raising conditions

These functions wrap the abort(), warn() and inform() functions from [rlang](https://rlang.r-lib.org/reference/abort.html), and apply default styling to messages.
```{r}
cli_abort("This is an Error!")
cli_warn("This is a Warning")
cli_inform("This is a Message")
```

For longer messages with multiple lines, you can pass in a named vector. Each element of the vector is a new line of text, with the names specifying (optional) bullet style:

```{r}
cli_abort(c("This is an Error.",
  "i" = "This is more information about the Error",
  " " = "continuing onto the next line"))
```

The first element of a vector passed to `cli_abort()` has the `"!"` bullet attached by default. 

<!--- wrap lines --->

The following [`cli_bullets()`](https://cli.r-lib.org/reference/cli_bullets.html) are available by default:

```{r}
cli_bullets(c(
        "noindent",
  " " = "indent",
  "*" = "bullet",
  ">" = "arrow",
  "v" = "success",
  "x" = "danger",
  "!" = "warning",
  "i" = "info"
))
```

## Inline Markup

<!--- https://cli.r-lib.org/reference/inline-markup.html#classes -->

Text within a line can be styled according to classes defined in the default theme [`builtin_theme()`](https://cli.r-lib.org/reference/builtin_theme.html):

```{r}
cli_bullets(c(
  "A Function Argument: {.arg path}",
  "A Package Name: {.pkg cli}",
  "A URL: {.url https://example.com}."
)
)

```

## Command Substitution

<!---https://cli.r-lib.org/reference/inline-markup.html --->

## Pluralization

## Themes

<!--- https://cli.r-lib.org/reference/builtin_theme.html -->