---
title: "Data import with the tidyverse :: Cheatsheet"
description: " "
execute:
  eval: true
  output: false
  warning: true
---

```{r}
#| output: asis
#| echo: false
#| column: margin
source("common.R")
use_cheatsheet_logo(
  "readr", 
  alt = "Hex logo for readr - a white document icon going dowards into a dark blue funnel, with a white icon of a data table coming out the narrow end, all on a blue background."
  )
sheet_name <- tools::file_path_sans_ext(knitr::current_input())
pdf_preview_link(sheet_name)
translation_list(sheet_name)
```

One of the first steps of a project is to import outside data into R.
Data is often stored in tabular formats, like csv files or spreadsheets.

-   The first half of this cheatsheet shows how to import and save text files into R using **readr**.
-   The second half shows how to import spreadsheet data from Excel files using **readxl** or Google Sheets using **googlesheets4**.

```{r}
library(readr)
library(readxl)
library(googlesheets4)
```

For importing other types of data try one of the following packages:

-   **haven**: SPSS, Stata, and SAS files
-   **DBI**: databases
-   **jsonlite**: json
-   **xml2**: XML
-   **httr**: Web APIs
-   **rvest**: HTML (Web Scraping)
-   **readr::read_lines()**: text data

## Read Tabular Data with readr

```{r}
#| include: false
library(readr)
```

See `?read_delim`.

``` r
read_*(
  file, 
  col_names = TRUE, col_types = NULL, col_select = NULL, 
  show_col_types = TRUE
  id = NULL, locale, 
  n_max = Inf, skip = 0, guess_max = min(1000, n_max), 
  na = c("", "NA")
)
```

### Examples

-   Read files with any delimiter: `read_delim()`.
    If no delimiter is specified, it will automatically guess.

    -   If the file you want to import is the following:

        ```         
        A|B|C
        1|2|3
        4|5|NA
        ```

        ```{r}
        #| label: write-file-txt
        #| echo: false

        write_file("A|B|C\n1|2|3\n4|5|NA", file = "file.txt")
        ```

    -   Read it with `read_delim()` and it will look like the following when imported:

        ```{r}

        read_delim("file.txt", delim = "|", show_col_types = FALSE)
        ```

    -   To make `file.txt`, run:

        ```{r}
        #| ref.label: write-file-txt
        #| eval: false
        ```

-   Read a comma delimited file with period decimal marks: `read_csv()`.

    -   If the file you want to import is the following:

        ```         
        A,B,C
        1,2,3
        4,5,NA
        ```

        ```{r}
        #| label: write-file-csv
        #| echo: false

        write_file("A,B,C\n1,2,3\n4,5,NA", file = "file.csv")
        ```

    -   Read it with `read_csv()` and it will look like the following when imported:

        ```{r}
        read_csv("file.csv", show_col_types = FALSE)
        ```

    -   To make `file.csv`, run:

        ```{r}
        #| ref.label: write-file-csv
        #| eval: false
        ```

-   Read semicolon delimited files with comma decimal marks: `read_csv2()`.

    -   If the file you want to import is the following:

        ```         
        A;B;C
        1,5;2;3
        4,5;5;NA
        ```

        ```{r}
        #| label: write-file-csv2
        #| echo: false

        write_file("A;B;C\n1,5;2;3\n4,5;5;NA", file = "file2.csv")
        ```

    -   Read it with `read_csv2()` and it will look like the following when imported:

        ```{r}
        read_csv2("file2.csv", show_col_types = FALSE)
        ```

    -   To make `file2.csv`, run:

        ```{r}
        #| ref.label: write-file-csv2
        #| eval: false
        ```

-   Read a tab delimited file: `read_tsv()` or `read_table()`.

    Read a fixed width file: `read_fwf("file.tsv", fwf_widths(c(2, 2, NA)))`.

    -   If the file you want to import is the following:

        ```         
        A B C
        1 2 3
        4 5 NA
        ```

        ```{r}
        #| label: write-file-tsv
        #| echo: false

        write_file("A\tB\tC\n1\t2\t3\n4\t5\tNA\n", file = "file.tsv")
        ```

    -   Read it with `read_tsv()` and it will look like the following when imported:

        ```{r}
        read_tsv("file.tsv", show_col_types = FALSE)
        ```

    -   To make `tsv`, run:

        ```{r}
        #| ref.label: write-file-tsv
        #| eval: false
        ```

### Useful read arguments

Suppose you have the following CSV files that you want to read in, called `file.csv`:

`file.csv`

```         
A,B,C
1,2,3
4,5,NA
```

`file3.csv`

```         
A,B,C
7,8,9
NA,11,12
```

To make these files, run:

```{r}
write_file("A,B,C\n1,2,3\n4,5,NA", file = "file.csv")
write_file("A,B,C\n7,8,9\nNA,11,12", file = "file3.csv")
```

-   No header: `col_names = FALSE`

    ```{r}
    #| message: false

    read_csv("file.csv", col_names = FALSE)
    ```

-   Provide header: `col_names = c("x", "y", "z")`

    ```{r}
    #| message: false

    read_csv("file.csv", col_names = c("x", "y", "z"))
    ```

-   Skip lines:

    ```{r}
    #| message: false

    read_csv("file.csv", skip = 1)
    ```

-   Read a subset of lines:

    ```{r}
    #| message: false

    read_csv("file.csv", n_max = 1)
    ```

-   Read values as missing:

    ```{r}
    #| message: false

    read_csv("file.csv", na = c("1"))
    ```

-   Specify decimal marks:

    ```{r}
    #| message: false

    read_delim("file2.csv", locale = locale(decimal_mark = ","))
    ```

-   Read multiple files into a single table:

    ```{r}
    #| message: false

    read_csv(c("file.csv", "file3.csv"), id = "origin_file")
    ```

### Save data with readr

``` r
write_*(
  x, file, 
  na = "NA", 
  append, col_names, quote, escape, eol, num_threads, progress
)
```

-   Write files with any delimiter: `write_delim(x, file, delim = " ")`
-   Write a comma delimited file: `write_csv(x, file)`
-   Write a semicolon delimited file: `write_csv2(x, file)`
-   Write a tab delimited file: `write_tsv(x, file)`

### Column specification with readr

Column specifications define what data type each column of a file will be imported as.
By default readr will generate a column spec when a file is read and output a summary.

`spec(df)`: Extract the full column specification for the given imported data frame.

```{r}
#| eval: false

spec(df)
# cols(
#   age = col_integer(),   # age is an integer
#   edu = col_character(), # edu is a character
#   earn = col_double()    # earn is a double (numeric)
# )
```

#### Column types

Each column type has a function and corresponding string abbreviation.

-   `col_logical() - "l"`
-   `col_integer() - "i"`
-   `col_double() - "d"`
-   `col_number() - "n"`
-   `col_character() - "c"`
-   `col_factor(levels, ordered = FALSE) - "f"`
-   `col_datetime(format = "") - "T"`
-   `col_date(format = "") - "D"`
-   `col_time(format = "") - "t"`
-   `col_skip() - "-", "_"`
-   `col_guess() - "?"`

#### Useful column arguments

-   Hide col spec message:

    ``` r
    read_*(file, show_col_types = FALSE)
    ```

-   Select columns to import: Use names, position, or selection helpers.

    ``` r
    read_*(file, col_select = c(age, earn))
    ```

-   Guess column types: To guess a column type, `read_ *()` looks at the first 1000 rows of data.
    Increase with `guess_max`.

    ``` r
    read_*(file, guess_max = Inf)
    ```

#### Define column specification

-   Set a default type:

    ```{r}
    #| eval: false

    read_csv(
      file, 
      col_type = list(.default = col_double())
    )
    ```

-   Use column type or string abbreviation:

    ```{r}
    #| eval: false

    read_csv(
      file, 
      col_type = list(x = col_double(), y = "l", z = "_")
    )
    ```

-   Use a single string of abbreviations:

    ```{r}
    #| eval: false

    # col types: skip, guess, integer, logical, character
    read_csv(
      file, 
      col_type = "_?ilc"
    )
    ```

## Import spreadsheets with readxl

### Read Excel files

Read a .xls or .xlsx file based on the file extension, e.g. `read_excel("excel_file.xlsx")`.
See [Useful read arguments] for more read arguments.
Also `read_xls()` and `read_xlsx()`.

```{r}
#| eval: false

read_excel(path, sheet = NULL, range = NULL)
```

-   If the Google sheet you want to import is the following:

    | A   | B   | C   | D   | E   |
    |-----|-----|-----|-----|-----|
    | x1  | x2  | x3  | x4  | x5  |
    | x   |     | z   | 8   |     |
    | y   | 7   |     | 9   | 10  |

    : Spreadsheet with 5 columns (A through E) and three rows. First row reads x1 through x5. Second and third row have some missing values.

-   It will look like the following when imported:

    ```{r}
    #| echo: false

    write_file("x1,x2,x3,x4,x5\nx, ,z,8, \ny,7, ,9,10", file = "excel.csv")
    read_csv("excel.csv", show_col_types = FALSE)
    ```

### Read sheets

-   Specify which sheet to read by position or name: `read_excel(path, sheet = NULL)`

    -   `read_excel(path, sheet = 1)`
    -   `read_excel(path, sheet = "s1")`

-   Get a vector of sheet names: `excel_sheets(path)`

    `excel_sheets("excel_file.xlsx")`

-   To read multiple sheets:

    1.  Get a vector of sheet names from the file path.

    2.  Set the vector names to be the sheet names.

    3.  Use `purrr::map()` and `purrr::list_rbind()` to read multiple files into one data frame.

        ```{r}
        #| eval: false

        path <- "your_file_path.xlsx"
        path |> 
          excel_sheets() |>
          set_names() |>
          map(read_excel, path = path) |>
          list_rbind()
        ```

### readxl column specification

-   Column specifications define what data type each column of a file will be imported as.

-   Use the `col_types` argument of `read_excel()` to set the column specification.

-   Guess column types: To guess a column type, `read_excel()` looks at the first 1000 rows of data.
    Increase with the `guess_max` argument.

    ```{r}
    #| eval: false

    read_excel(path, guess_max = Inf)
    ```

-   Set all columns to same type, e.g. character:

    ```{r}
    #| eval: false

    read_excel(path, col_types = "text")
    ```

-   Set each column individually:

    ```{r}
    #| eval: false

    read_excel(
      path,
      col_types = c("text", "guess", "guess","numeric")
    )
    ```

-   **Column types:**

    | logical | numeric | text  | date       | list  |
    |---------|---------|-------|------------|-------|
    | TRUE    | 2       | hello | 1947-01-08 | hello |
    | FALSE   | 3.45    | world | 1956-10-21 | 1     |

    : Table with 5 columns. Column headers are various data types (logical, numeric, text, date, and list). The data in two rows show examples of data for the given column type.

    -   `skip`

    -   `guess`

    -   `logical`

    -   `date`

    -   `numeric`

    -   `text`

    -   Use `list`for columns that include multiple data types.
        See **tidyr** and **purrr** for list-column data.

### Other useful Excel packages

-   For functions to write data to Excel files: **openxlsx** and **writexl**
-   For working with non-tabular Excel data: **tidyxl**

## Import spreadsheets with googlesheets4

### Read sheets

Read a sheet from a URL, a Sheet ID, or a dribble samefrom the googledrive package.
See [Useful read arguments] for more read arguments.

```{r}
#| eval: false

read_sheet(ss, sheet = NULL, range = NULL)
```

Same as `range_read()`.

-   If the Google sheet you want to import is the following:

    | A   | B   | C   | D   | E   |
    |-----|-----|-----|-----|-----|
    | x1  | x2  | x3  | x4  | x5  |
    | x   |     | z   | 8   |     |
    | y   | 7   |     | 9   | 10  |

    : Spreadsheet with 5 columns (A through E) and three rows. First row reads x1 through x5. Second and third row have some missing values.

-   It will look like the following when imported:

    ```{r}
    #| echo: false

    write_file("x1,x2,x3,x4,x5\nx, ,z,8, \ny,7, ,9,10", file = "googlesheet.csv")
    read_csv("googlesheet.csv", show_col_types = FALSE)
    ```

### Sheet metadata

-   **URLs** are in the form:

    ```         
    https://docs.google.com/spreadsheets/d/
                SPREADSHEET_ID/edit#gid=SHEET_ID
    ```

-   Get spreadsheet meta data: `gs4_get(ss)`

-   Get data on all spreadsheet files: `gs4_find(...)`

-   Get a tibble of properties for each worksheet: `sheet_properties(ss)`.
    Also `sheet_names()`.

### Write sheets

-   `write_sheet(data, ss = NULL, sheet = NULL)`**:** Write a data frame into a new or existing Sheet. 
-   `gs4_create(name, ..., sheets = NULL)`: Create a new Sheet with a vector of names, a data frame, or a (named) list of data frames.
-   `sheet_append(ss, data, sheet = 1)`: Add rows to the end of a worksheet.

### googlesheets4 column specification

Column specifications define what data type each column of a file will be imported as.

Use the `col_types` argument of `read_sheet()`**/**`range_read()` to set the column specification.

-   Guess column types: To guess a column type `read_sheet()`/`range_read()` looks at the first 1000 rows of data.
    Increase with `guess_max`.

    ```{r}
    #| eval: false

    read_sheet(path, guess_max = Inf)
    ```

-   Set all columns to same type, e.g. character:

    ```{r}
    #| eval: false

    read_sheet(path, col_types = "c")
    ```

-   Set each column individually:

    ```{r}
    #| eval: false

    # col types: skip, guess, integer, logical, character
    read_sheets(ss, col_types = "_?ilc")
    ```

-   Column types:

    -   skipped my lunch 🥙 🍱 and: "\_" or "-"

    -   guess: "?"

    -   logical: "l"

    -   integer: "i"

    -   double: "d"

    -   numeric: "n"

    -   date: "D"

    -   datetime: "T"

    -   character: "c"

    -   list-column: "L"

    -   cell: "C" (returns list of raw cell data)

    -   Use list for columns that include multiple data types.
        See **tidyr** and **purrr** for list-column data.

### File level operations

-   **googlesheets4** also offers ways to modify other aspects of Sheets (e.g. freeze rows, set column width, manage (work)sheets). Go to [googlesheets4.tidyverse.org](https://googlesheets4.tidyverse.org/) to read more.
-   For whole-file operations (e.g. renaming, sharing, placing within a folder), see the tidyverse package **googledrive** at [googledrive.tidyverse.org](https://googledrive.tidyverse.org).

## Cell specification for readxl and googlesheets4

Use the **range** argument of **readxl::read_excel()** or **googlesheets4::read_sheet()** to read a subset of cells from a sheet.

```{r}
#| eval: false

read_excel(path, range = "Sheet1!B1:D2")
read_sheet(ss, range = "B1:D2")
```

Also use the range argument with cell specification functions `cell_limits()`, `cell_rows()`, `cell_cols()`, and `anchored()`**.**

------------------------------------------------------------------------

CC BY SA Posit Software, PBC • [info\@posit.co](mailto:info@posit.co) • [posit.co](https://posit.co)

Learn more at

-   readr: [readr.tidyverse.org](https://readr.tidyverse.org)
-   readxl: [readxl.tidyverse.org](https://readxl.tidyverse.org/)
-   googlesheets4: [googlesheets4.tidyverse.org](https://googlesheets4.tidyverse.org/)

Updated: `r format(Sys.Date(), "%Y-%m")`.

```{r}
#| output: true

packageVersion("readr")
packageVersion("readxl")
packageVersion("googlesheets4")
```

------------------------------------------------------------------------

```{r include = FALSE}
fs::file_delete(
  c("file.txt",
  "file.csv",
  "file2.csv",
  "file3.csv",
  "file.tsv",
  "excel.csv",
  "googlesheet.csv")
)
```
