---
title: "Factors with forcats :: Cheatsheet"
description: " "
image-alt: ""
execute:
  eval: true
  output: false
  warning: false
---

```{r}
#| output: asis
#| echo: false
#| column: margin
source("common.R")
use_cheatsheet_logo(
  "forcats", 
  alt = "Hex logo for forcats - drawing of four black cats lounging in a cardboard box. On one side of the box it says 'for' and on the adjacent side is says 'cats'."
)
sheet_name <- tools::file_path_sans_ext(knitr::current_input())
pdf_preview_link(sheet_name)
translation_list(sheet_name)
```

The **forcats** package provides tools for working with factors, which are R's data structure for categorical data.

```{r}
library(forcats)
```

```{r}
#| label: other-used-packages
#| include: false

library(ggplot2)
```

## Factors

R represents categorical data with factors.
A **factor** is an integer vector with a **levels** attribute that stores a set of mappings between integers and categorical values.
When you view a factor, R displays not the integers but the levels associated with them.

For example, R will display `c("a", "c", "b", "a")` with levels `c("a", "b", "c")` but will store `c(1, 3, 2, 1)` where 1 = a, 2 = b, and 3 = c.

R will display:

```{r}
#| echo: false
#| output: true

f <- factor(c("a", "c", "b", "a"), levels = c("a", "b", "c"))
f
```

R will store:

```{r}
#| echo: false
#| output: true

unclass(f)
```

Create a factor with `factor()`:

-   `factor(x = character(), levels, labels = levels, exclude = NA, ordered = is.ordered(x), nmax = NA)`: Convert a vector to a factor.
    Also `as_factor()`.

    ```{r}
    f <- factor(c("a", "c", "b", "a"), levels = c("a", "b", "c"))
    ```

Return its levels with `levels()`:

-   `levels(x)`: Return/set the levels of a factor.

    ```{r}
    levels(f)
    levels(f) <- c("x", "y", "z")
    ```

Use `unclass()` to see its structure.

## Inspect Factors

-   `fct_count(f, sort = FALSE, prop = FALSE)`: Count the number of values with each level.

    ```{r}
    fct_count(f)
    ```

-   `fct_match(f, lvls)`: Check for `lvls` in `f`.

    ```{r}
    #| include: false

    f <- factor(c("a", "c", "b", "a"), levels = c("a", "b", "c"))
    ```

    ```{r}
    fct_match(f, "a")
    ```

-   `fct_unique(f)`: Return the unique values, removing duplicates.

    ```{r}
    fct_unique(f)
    ```

## Combine Factors

-   `fct_c(...)`: Combine factors with different levels.
    Also `fct_cross()`.

    ```{r}
    f1 <- factor(c("a", "c"))
    f2 <- factor(c("b", "a"))
    fct_c(f1, f2)
    ```

-   `fct_unify(fs, levels = lvls_union(fs))`: Standardize levels across a list of factors.

    ```{r}
    fct_unify(list(f2, f1))
    ```

## Change the order of levels

-   `fct_relevel(.f, ..., after = 0L)`: Manually reorder factor levels.

    ```{r}
    fct_relevel(f, c("b", "c", "a"))
    ```

-   `fct_infreq(f, ordered = NA)`: Reorder levels by the frequency in which they appear in the data (highest frequency first).
    Also `fct_inseq()`.

    ```{r}
    f3 <- factor(c("c", "c", "a"))
    fct_infreq(f3)
    ```

-   `fct_inorder(f, ordered = NA)`: Reorder levels by order in which they appear in the data.

    ```{r}
    fct_inorder(f2)
    ```

-   `fct_rev(f)`: Reverse level order.

    ```{r}
    f4 <- factor(c("a","b","c"))
    fct_rev(f4)
    ```

-   `fct_shift(f)`: Shift levels to left or right, wrapping around end.

    ```{r}
    fct_shift(f4)
    ```

-   `fct_shuffle(f, n = 1L)`: Randomly permute order of factor levels.

    ```{r}
    fct_shuffle(f4)
    ```

-   `fct_reorder(.f, .x, .fun = median, ..., .desc = FALSE)`: Reorder levels by their relationship with another variable.

    ```{r}
    boxplot(PlantGrowth, weight ~ fct_reorder(group, weight))
    ```

-   `fct_reorder2(.f, .x, .y, .fun = last2, ..., .desc = TRUE)`: Reorder levels by their final values when plotted with two other variables.

    ```{r}
    ggplot(
      diamonds,
      aes(carat, price, color = fct_reorder2(color, carat, price))
      ) + 
      geom_smooth()
    ```

## Change the value of levels

-   `fct_recode(.f, ...)`: Manually change levels.
    Also `fct_relabel()` which obeys `purrr::map` syntax to apply a function or expression to each level.

    ```{r}
    fct_recode(f, v = "a", x = "b", z = "c")
    fct_relabel(f, ~ paste0("x", .x))
    ```

-   `fct_anon(f, prefix = "")`: Anonymize levels with random integers.

    ```{r}
    fct_anon(f)
    ```

-   `fct_collapse(.f, …, other_level = NULL)`: Collapse levels into manually defined groups.

    ```{r}
    fct_collapse(f, x = c("a", "b"))
    ```

-   `fct_lump_min(f, min, w = NULL, other_level = "Other")`: Lumps together factors that appear fewer than `min` times.
    Also `fct_lump_n()`, `fct_lump_prop()`, and `fct_lump_lowfreq()`.

    ```{r}
    fct_lump_min(f, min = 2)
    ```

-   `fct_other(f, keep, drop, other_level = "Other")`: Replace levels with "other."

    ```{r}
    fct_other(f, keep = c("a", "b"))
    ```

## Add or drop levels

-   `fct_drop(f, only)`: Drop unused levels.

    ```{r}
    f5 <- factor(c("a","b"),c("a","b","x"))
    f6 <- fct_drop(f5)
    ```

-   `fct_expand(f, ...)`: Add levels to a factor.

    ```{r}
    fct_expand(f6, "x")
    ```

-   `fct_na_value_to_level(f, level = "(Missing)")`: Assigns a level to NAs to ensure they appear in plots, etc.

    ```{r}
    f <- factor(c("a", "b", NA))
    fct_na_value_to_level(f, level = "(Missing)")
    ```

------------------------------------------------------------------------

CC BY SA Posit Software, PBC • [info\@posit.co](mailto:info@posit.co) • [posit.co](https://posit.co)

Learn more at [forcats.tidyverse.org](https://forcats.tidyverse.org).

Updated: `r format(Sys.Date(), "%Y-%m")`.

```{r}
#| output: true

packageVersion("forcats")
```

------------------------------------------------------------------------
