---
title: "Dates and times with lubridate :: Cheatsheet"
description: " "
image-alt: ""
execute:
  eval: true
  output: false
  warning: false
---

```{r}
#| output: asis
#| echo: false
#| column: margin
source("common.R")
use_cheatsheet_logo(
  "lubridate", 
  alt = "Hex logo for lubridate - a white calendar on a green background, with a clock superimposed on it. 'lubdridate' is written in white across the bottom.")
sheet_name <- tools::file_path_sans_ext(knitr::current_input())
pdf_preview_link(sheet_name)
translation_list(sheet_name)
```

## Date-times

```{r}
library(lubridate)
```

```{r}
#| include: false

library(hms)
```

A **date-time** is a point on the timeline, stored as the number of seconds since 1970-01-01 00:00:00 UTC

```{r}
dt <- as_datetime(1511870400)
```

A **date** is a day stored as the number of days since 1970-01-01

```{r}
d <- as_date(17498)
```

An hms is a **time** stored as the number of seconds since 00:00:00

```{r}
t <- hms::as_hms(85)
```

### Parse Date-Times

Convert strings or numbers to date-times

1.  Identify the order of the year (**y**), month (**m**), day (**d**), hour (**h**), minute (**m**) and second (**s**) elements in your data.

2.  Use the function below whose name replicates the order.
    Each accepts a `tz` argument to set the time zone, e.g. `ymd(x, tz = "UTC")`.

-   `ymd_hms()`, `ymd_hm()`, `ymd_h()`

    ```{r}
    ymd_hms("2017-11-28T14:02:00")
    ```

-   `ydm_hms()`, `ydm_hm()`, `ydm_h()`

    ```{r}
    ydm_hms("2017-22-12 10:00:00")
    ```

-   `mdy_hms()`, `mdy_hm()`, `mdy_h()`

    ```{r}
    mdy_hms("11/28/2017 1:02:03")
    ```

-   `dmy_hms()`, `dmy_hm()`, `dmy_h()`

    ```{r}
    dmy_hms("1 Jan 2017 23:59:59")
    ```

-   `ymd()`, `ydm()`

    ```{r}
    ymd(20170131)
    ```

-   `mdy()`, `myd()`

    ```{r}
    mdy("July 4th, 2000")
    ```

-   `dmy()`, `dym()`

    ```{r}
    dmy("4th of July '99")
    ```

-   `yq()`: Q for quarter.

    ```{r}
    yq("2001: Q3")
    ```

-   `my()`, `ym()`

    ```{r}
    my("07-2020")
    ```

-   `hms::hms()`: Also `lubridate::hms()`, `hm()`, and `ms()`, which return periods\*.

    ```{r}
    hms::hms(seconds = 0, minutes = 1, hours = 2)
    ```

-   `date_decimal(decimal, tz = "UTC")`

    ```{r}
    date_decimal(2017.5)
    ```

-   `now(tzone = "")`: Current time in tz (defaults to system tz).

    ```{r}
    now()
    ```

-   `today(tzone = "")`: Current date in a tz (defaults to system tz).

    ```{r}
    today()
    ```

-   `fast_strptime()`: Faster strptime.

    ```{r}
    fast_strptime("9/1/01", "%y/%m/%d")
    ```

-   `parse_date_time()`: Easier strptime.

    ```{r}
    parse_date_time("09-01-01", "ymd")
    ```

### Get and Set Components

Use an accessor function to get a component.

```{r}
#| output: true

d
day(d)
```

Assign into an accessor function to change a component in place.

```{r}
#| output: true

day(d) <- 1
d
```

-   `date(x)`: Date component.

    ```{r}
    date(dt)
    ```

-   `year(x)`: Year.

    ```{r}
    year(dt)
    ```

-   `isoyear(x)`: The ISO 8601 year.

-   `epiyear(x)`: Epidemiological year.

-   `month(x, label, abbr)`: Month.

    ```{r}
    month(dt)
    ```

-   `day(x)`: Day of the month.

    ```{r}
    day(dt)
    ```

-   `wday(x, label, abbr)`: Day of week.

-   `qday(x)`: Day of quarter.

-   `hour(x)`: Hour.

    ```{r}
    hour(dt)
    ```

-   `minute(x)`: Minutes.

    ```{r}
    minute(dt)
    ```

-   `second(x)`: Seconds.

    ```{r}
    second(dt)
    ```

-   `tz(x)`: Time zone.

    ```{r}
    tz(dt)
    ```

-   `week(x)`: Week of the year.

    ```{r}
    week(dt)
    ```

-   `isoweek()`: ISO 8601 week.

-   `epiweek()`: Epidemiological week.

-   `quarter(x)`: Quarter.

    ```{r}
    quarter(dt)
    ```

-   `semester(x, with_year = FALSE)`: Semester.

    ```{r}
    semester(dt)
    ```

-   `am(x)`: Is it in the am?

    ```{r}
    am(dt)
    ```

-   `pm(x)`: Is it in the pm?

    ```{r}
    pm(dt)
    ```

-   `dst(x)`: Is it daylight savings?

    ```{r}
    dst(d)
    ```

-   `leap_year(x)`: Is it a leap year?

    ```{r}
    leap_year(d)
    ```

-   `update(object, ..., simple = FALSE)`

    ```{r}
    update(dt, mday = 2, hour = 1)
    ```

## Round Date-times

-   `floor_date(x, unit = "second")`: Round down to nearest unit.

    ```{r}
    floor_date(dt, unit = "month")
    ```

-   `round_date(x, unit = "second")`: Round to nearest unit.

    ```{r}
    round_date(dt, unit = "month")
    ```

-   `ceiling_date(x, unit = "second")`: Round up to the nearest unit.

    ```{r}
    ceiling_date(dt, unit = "month")
    ```

Valid units are second, minute, hour, day, week, month, bimonth, quarter, season, halfyear and year.

-   `rollback(dates, roll_to_first = FALSE, preserve_hms = TRUE)`: Roll back to last day of previous month.
    Also `rollforward()`.

    ```{r}
    rollback(dt)
    ```

## Stamp Date-times

`stamp()`: Derive a template from an example string and return a new function that will apply the template to date-times.
Also `stamp_date()` and `stamp_time()`.

1.  Derive a template, create a function

    ```{r}
    sf <- stamp("Created Sunday, Jan 17, 1999 3:34")
    ```

2.  Apply the template to dates

    ```{r}
    #| output: true

    sf(ymd("2010-04-05"))
    ```

**Tip: use a date with day \> 12**

## Time Zones

R recognizes \~600 time zones.
Each encodes the time zone, Daylight Savings Time, and historical calendar variations for an area.
R assigns *one* time zone per vector.

Use the `UTC` time zone to avoid Daylight Savings.

-   `OlsonNames()`: Returns a list of valid time zone names.

    ```{r}
    OlsonNames()
    ```

-   `Sys.timezone()`: Gets current time zone.

-   `with_tz(time, tzone = "")`: Get the **same date-time** in a new time zone (a new clock time).
    Also `local_time(dt, tz, units)`.
    For example, 4:00 Pacific becomes 5:00 Mountain, or 6:00 Central, or 7:00 Eastern.

    ```{r}
    with_tz(dt, "US/Pacific")
    ```

-   `force_tz(time, tzone = "")`: Get the **same clock time** in a new time zone (a new date-time).
    Also `force_tzs()`.
    For example, 7:00 Pacific becomes 7:00 Mountain, or 7:00 Central, or 7:00 Eastern.

    ```{r}
    force_tz(dt, "US/Pacific")
    ```

## Math with Date-times

Lubridate provides three classes of timespans to facilitate math with dates and date-times.

Math with date-times relies on the **timeline**, which behaves inconsistently.
Consider how the timeline behaves during:

-   A normal day:

```{r}
nor <- ymd_hms("2018-01-01 01:30:00", tz = "US/Eastern")
```

-   The start of daylight savings (spring forward):

```{r}
gap <- ymd_hms("2018-03-11 01:30:00", tz = "US/Eastern")
```

-   The end of daylight savings (fall back):

```{r}
lap <- ymd_hms("2018-11-04 00:30:00", tz = "US/Eastern")
```

-   Leap years and leap seconds:

```{r}
leap <- ymd("2019-03-01")
```

**Periods** track changes in clock times, which ignore time line irregularities.

```{r}
nor + minutes(90)
gap + minutes(90)
lap + minutes(90)
leap + years(1)
```

**Durations** track the passage of physical time, which deviates from clock time when irregularities occur.

```{r}
nor + dminutes(90)
gap + dminutes(90)
lap + dminutes(90)
leap + dyears(1)
```

**Intervals** represent specific intervals of the timeline, bounded by start and end date-times.

```{r}
interval(nor, nor + minutes(90))
interval(gap, gap + minutes(90))
interval(lap, lap + minutes(90))
interval(leap, leap + years(1))
```

Not all years are 365 days due to **leap days**.
Not all minutes are 60 seconds due to **leap seconds**.
It is possible to create an imaginary date by adding **months**, e.g.
February 31st.

```{r}
#| output: true

jan31 <- ymd(20180131)
jan31 + months(1)
```

`%m+%` and `%m-%` will roll imaginary dates to the last day of the previous month.

```{r}
#| output: true

jan31 %m+% months(1)
```

`add_with_rollback(e1, e2, roll_to_first = TRUE)` will roll imaginary dates to the first day of the new month.

```{r}
#| output: true

add_with_rollback(jan31, months(1), roll_to_first = TRUE)
```

### Periods

Add or subtract periods to model events that happen at specific clock times, like the NYSE opening bell.

Make a period with the name of a time unit **pluralized**, e.g.

```{r}
#| output: true

p <- months(3) + days(12)

# Shows the number of months, number of days, etc.
p
```

-   `years(x = 1)`: x years.

-   `months(x = 1)`: x months.

-   `weeks(x = 1)`: x weeks.

-   `days(x = 1)`: x days.

-   `hours(x = 1)`: x hours.

-   `minutes(x = 1)`: x minutes.

-   `seconds(x = 1)`: x seconds.

-   `milliseconds(x = 1)`: x milliseconds.

-   `microseconds(x = 1)`: x microseconds.

-   `nanoseconds(x = 1)`: x nanoseconds.

-   `picoseconds(x = 1)`: x picoseconds.

-   `period(num = NULL, units = "second", ...)`: An automation friendly period constructor.

    ```{r}
    period(5, unit = "years")
    ```

-   `as.period(x, unit)`: Coerce a timespan to a period, optionally in the specified units.
    Also `is.period()`.

    ```{r}
    as.period(p)
    ```

-   `period_to_seconds(x)`: Convert a period to the "standard" number of seconds implied by the period.
    Also `seconds_to_period()`.

    ```{r}
    period_to_seconds(p)
    ```

### Durations

Add or subtract durations to model physical processes, like battery life.
Durations are stored as seconds, the only time unit with a consistent length.
**Difftimes** are a class of durations found in base R.

Make a duration with the name of a period prefixed with a *d*, e.g.

```{r}
#| output: true

dd <- ddays(14)

# Shows the exact length in seconds, and the equivalent in common units
dd
```

-   `dyears(x = 1)`: 31536000x seconds.

-   `dmonths(x = 1)`: 2629800x seconds.

-   `dweeks(x = 1)`: 604800x seconds.

-   `ddays(x = 1)`: 86400x seconds.

-   `dhours(x = 1)`: 3600x seconds.

-   `dminutes(x = 1)`: 60x seconds.

-   `dseconds(x = 1)`: x seconds.

-   `dmilliseconds(x = 1)`: x \* 10^-3^ seconds.

-   `dmicroseconds(x = 1)`: x \* 10^-6^ seconds.

-   `dnanoseconds(x = 1)`: x \* 10^-9^ seconds.

-   `dpicoseconds(x = 1)`: x \* 10^-12^ seconds.

-   `duration(num = NULL, units = "second", ...)`: An automation friendly duration constructor.

    ```{r}
    duration(5, unit = "years")
    ```

-   `as.duration(x, ...)`: Coerce a timespan to a duration.
    Also `is.duration()`, `is.difftime()`.

    ```{r}
    as.duration(p)
    ```

-   `make_difftime(x)`: Make diffime with the specified number of units.

    ```{r}
    make_difftime(99999)
    ```

### Intervals

Divide an interval by a duration to determine its physical length, divide by an interval by a period to determine its implied length in clock time.

Make an interval with `interval()` or `%--%`, e.g.

```{r}
#| output: true

i <- interval(ymd("2017-01-01"), d)
i

# Shows the exact length in seconds, and the equivalent in common units
j <- d %--% ymd("2017-12-31")
j
```

-   `a %within% b`: Does interval or dte0time `a` fall within interval `b`?

    ```{r}
    now() %within% i
    ```

-   `int_start(int)`: Access/set the start date-time of an interval.
    Also `int_end()`.

    ```{r}
    int_start(i) <- now()
    int_start(i)
    ```

-   `int_aligns(int1, int2)`: Do two intervals share a boundary?
    Also `int_overlaps()`.

    ```{r}
    int_aligns(i, j)
    ```

-   `int_diff(times)`: Make the intervals that occur between the date-times in a vector.

    ```{r}
    v <- c(dt, dt + 100, dt + 1000)
    int_diff(v)
    ```

-   `int_flip(int)`: Reverse the direction of an interval.
    Also `int_standardize()`.

    ```{r}
    int_flip(i)
    ```

-   `int_length(int)`: Length in seconds.

    ```{r}
    int_length(i)
    ```

-   `int_shift(int, by)`: Shifts an interval up or down the timeline by a timespan.

    ```{r}
    int_shift(i, days(-1))
    ```

-   `as.interval(x, start, ...)`: Coerce a timespan to an interval with the start date-time.
    Also `is.interval()`.

    ```{r}
    as.interval(days(-1), start = now())
    ```

------------------------------------------------------------------------

CC BY SA Posit Software, PBC • [info\@posit.co](mailto:info@posit.co) • [posit.co](https://posit.co)

Learn more at [lubridate.tidyverse.org](https://lubridate.tidyverse.org).

Updated: `r format(Sys.Date(), "%Y-%m")`.

```{r}
#| output: true

packageVersion("lubridate")
```

------------------------------------------------------------------------
