---
title: "Create models with parsnip :: Cheatsheet"
description: " "
image-alt: ""
execute:
  eval: true
  warning: false
---

```{css}
#| echo: false
.package-pills {
  display: flex;
  flex-wrap: wrap;
  gap: 8px;
  margin: 4px 0 10px 0;
}

.pill {
  display: inline-block;
  padding: 2px 12px;
  margin: 2px 0;
  border: 1px solid #6c757d;
  border-radius: 20px;
  font-size: 14px;
  font-family: monospace;
  color: #333;
  background-color: white;
  white-space: nowrap;
}

.pill sup {
  font-size: 10px;
  margin-left: 3px;
}
```

```{r}
#| output: asis
#| echo: false
#| column: margin
source("common.R")
use_cheatsheet_logo(
  "parsnip", 
  alt = "Hex logo for recipes, drawing of a cartoon parsnip smiling."
)
sheet_name <- tools::file_path_sans_ext(knitr::current_input())
pdf_preview_link(sheet_name)
translation_list(sheet_name)

library(tidymodels)

package_pills <- function(packages) {
  cat('::: {.package-pills}\n')
  for (pkg in packages) {
    name <- pkg[[1]]
    numbers <- pkg[[2]]
    sups <- paste0('<sup>', numbers, '</sup>', collapse = ' ')
    cat('<span class="pill">', name, ' ', sups, '</span>\n', sep = '')
  }
  cat(':::\n')
}
```

## Basics

`parsnip` provides a tidy, unified interface to a range of models from other packages. It helps avoid having to remember how to properly call the modeling functions of those external packages.

A `parsnip` specification is made up of 3 main components:

1.  The type of model to be used, such as Random Forest (`rand_forest()`) or linear regression (`linear_reg()`)

2.  How will the model be used, or mode. The two most common are “regression” and “classification”.

3.  The computational engine, or program that will actually execute the training. It could be an external R package, such as ranger, or even an engine outside of R, such as Stan or Apache Spark.

```{r}
library(tidymodels)

rand_forest(mtry = 10, trees = 2000) |> # Define type of model
  set_engine("ranger", importance = "impurity") |> # Select an engine
  set_mode("regression") # Set the mode
```

`set_engine(object, engine, ...)` - Specifies which package or system will be used to fit the model, along with any arguments specific to that software.

`set_args(object, ...)` - Modifies the arguments of a model specification

`set_mode(object, mode, …)` - Changes the model's mode.

`show_engines(x)` - The possible engines for a model can depend on what packages are loaded. Some parsnip extension add engines to existing models.

```{r}
show_engines("linear_reg")
```

## Classification Only

-   `logistic_reg(mode = "classification", engine = "glm", penalty, mixture)` - Generalized linear model for binary outcomes. A linear combination of the predictors is used to model the log odds of an event.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("brulee", 1),
      list("gee", 1),
      list("glm", 1),
      list("glmer", 1),
      list("glmnet", 1),
      list("h2o", 1),
      list("keras", 1),
      list("LiblineaR", 1),
      list("spark", 1),
      list("stan", 1),
      list("stan_glmer", 1)
    ))
    ```
-   `multinom_reg(mode = "classification", engine = "nnet", penalty, mixture)` - Uses linear predictors to predict multiclass data using the multinomial distribution.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("brulee", 1),
      list("glmnet", 1),
      list("h2o", 1),
      list("keras", 1),
      list("nnet", 1),
      list("spark", 1)
    ))
    ```

-   `naive_Bayes(mode = "classification", smoothness, Laplace, engine = "klaR")` - Uses Bayes' theorem to compute the probability of each class, given the predictor values.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("h2o", 1),
      list("klaR", 1),
      list("naivebayes", 1)
    ))
    ```

-   `null_model(mode = "classification", engine = "parsnip")` - Fit a single mean or largest class model. This is the user-facing function for the null_model() specification.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("parsnip", 1)
    ))
    ```

-   `ordinal_reg(mode = "classification", ordinal_link, odds_link, penalty, mixture, engine = "polr")` - Defines a generalized linear model that predicts an ordinal outcome.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("rpartScore", 1),
      list("polr", 1),
      list("vgam", 1),
      list("vglm", 1)
    ))
    ```

## Regression Only

-   `linear_reg(mode = "regression", engine = "lm", penalty, mixture)` - Defines a model that can predict numeric values from predictors using a linear function.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("brulee", 2),
      list("gee", 2),
      list("glm", 2),
      list("glmer", 2),
      list("glmnet", 2),
      list("gls", 2),
      list("h2o", 2),
      list("keras", 2),
      list("lm", 2),
      list("lme", 2),
      list("quantreg", 2),
      list("spark", 2),
      list("stan", 2),
      list("stan_glmer", 2)
    ))
    ```

-   `poisson_reg(mode = "regression", penalty, mixture, engine = "glm")` - Defines a generalized linear model for count data that follow a Poisson distribution.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("gee", 2),
      list("glm", 2),
      list("glmer", 2),
      list("glmnet", 2),
      list("h2o", 2),
      list("hurdle", 2),
      list("stan", 2),
      list("stan_glmer", 2),
      list("zeroinfl", 2)
    ))
    ```

## General Use

-   `decision_tree(mode, engine = "rpart", cost_complexity, tree_depth, min_n)` - A set of if/then statements creates a tree-based structure.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("partykit", 123),
      list("rpart", 123),
      list("spark", 12),
      list("C5.0", 1)
    ))
    ```

-   `mars(mode, engine = "earth", num_terms, prod_degree, prune_method)` - Uses artificial features for some predictors. These features resemble hinge functions and the result is a model that is a segmented regression in small dimensions.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("earth", 12)
    ))
    ```

-   `mlp(mode, engine = "nnet", hidden_units, penalty, dropout, epochs, activation, learn_rate)` - Defines a multilayer perceptron model (a.k.a. a single layer, feed-forward neural network).

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("nnet", 12),
      list("brulee", 12),
      list("brulee_two_layer", 12),
      list("keras", 12),
      list("grnn", 12)
    ))
    ```

-   `gen_additive_mod(mode, select_features, adjust_deg_free, engine = "mgcv")` - Uses smoothed functions of numeric predictors in a generalized linear model.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("mgcv", 12)
    ))
    ```

-   `nearest_neighbor(mode, engine = "kknn", neighbors, weight_func, dist_power)` - Uses the K most similar data points from the training set to predict new samples.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("knn", 12)
    ))
    ```

-   `pls(mode, predictor_prop, num_comp, engine = "mixOmics")` - Uses latent variables to model the data. Similar to a supervised version of PCA.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("mixOmics", 12)
    ))
    ```

## Discriminant

-   `discrim_flexible(mode = "classification", num_terms, prod_degree, prune_method, engine = "earth")` - Fits a discriminant analysis model that uses nonlinear features created using MARS.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("earth", 1)
    ))
    ```

-   `discrim_regularized(mode = "classification", frac_common_cov, frac_identity, engine = "klaR")` - Estimates a multivariate distribution for the predictors separately for the data in each class. The model's structure can be LDA, QDA, or a combination. Each probability class is computed using Bayes's theorem, given the predictor values.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("klaR", 1)
    ))
    ```


<div style="text-align: center;">
Estimates a multivariate distribution for the predictors separately for the data in each class using a method described below. Each class' probability is computed using Bayes' theorem, given the predictor values.
</div>

-   `discrim_linear(mode = "classification", regularization_method, engine = "MASS", penalty)` - Uses Gaussian with a common covariance matrix to perform the estimate.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("MASS", 1),
      list("mda", 1),
      list("sda", 1),
      list("sparsediscrim", 1)
    ))
    ```

-   `discrim_quad(mode = "classification", regularization_method, engine = "MASS")` - Uses Gaussian with separate covariance matrices to perform the estimate.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("MASS", 1),
      list("sparsediscrim", 1)
    ))
    ```

## Support Vector Machine

<div style="text-align: center;">
**Classification:** Maximizes the width of the margin between classes using a method described below.

**Regression:** Optimizes a robust loss function only affected by very large model residuals and uses an additional method described below.
</div>

-   `svm_linear(mode, cost, engine = "LiblineaR", margin)` - Classification: A linear class boundary. Regression: Uses a linear fit.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("kernlab", 12),
      list("LiblineaR", 12)
    ))
    ```

-   `svm_poly(mode, cost, engine = "kernlab", degree, scale_factor)` - Classification: A polynomial class boundary. Regression: Uses polynomial functions of the predictors.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("kernlab", 12)
    ))
    ```

-   `svm_rbf(mode, cost, engine = "kernlab", rbf_sigma)` - Classification: A nonlinear class boundary. Regression: Uses nonlinear functions of the predictors.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("kernlab", 12)
    ))
    ```

## Feature Rules

-   `rule_fit(mode, mtry, trees, min_n, tree_depth, learn_rate, loss_reduction, sample_size, stop_iter, penalty, engine = "xrf")` - Derives simple feature rules from a tree ensemble and uses them as features in a regularized model.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("xrf", 12),
      list("h2o", 1)
    ))
    ```

-   `C5_rules(mode = "classification", trees, min_n, engine = "C5.0")` - Derives feature rules from a tree for prediction. A single tree or boosted ensemble can be used.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("C5.0", 1)
    ))
    ```

-   `cubist_rules(mode = "regression", committees, neighbors, max_rules, engine = "Cubist")` - Derives simple feature rules from a tree ensemble and creates regression models within each rule.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("Cubist", 2)
    ))
    ```

