---
title: "Create models with parsnip :: Cheatsheet"
description: " "
image-alt: ""
execute:
  eval: true
  warning: false
---

```{r}
#| output: asis
#| echo: false
#| column: margin
source("common.R")
use_cheatsheet_logo(
  "parsnip", 
  alt = "Hex logo for recipes, drawing of a cartoon parsnip smiling."
)
sheet_name <- tools::file_path_sans_ext(knitr::current_input())
pdf_preview_link(sheet_name)
translation_list(sheet_name)

library(tidymodels)
```


## Basics

`parsnip` provides a tidy, unified interface to a range of models from other packages. It helps avoid having to remember how to properly call the modeling functions of those external packages.

A `parsnip` specification is made up of 3 main components:

1. The type of model to be used, such as Random Forest (`rand_forest()`) or linear regression (`linear_reg()`)

1. How will the model be used, or mode. The two most common are “regression” and “classification”.

1. The computational engine, or program that will actually execute the training. It could be an external R package, such as ranger, or even an engine outside of R, such as Stan or Apache Spark.


```{r}
library(tidymodels)

rand_forest(mtry = 10, trees = 2000) |> # Define type of model
  set_engine("ranger", importance = "impurity") |> # Select an engine
  set_mode("regression") # Set the mode
```


`set_engine(object, engine, ...)` - Specifies which package or system will be used to fit the model, along with any arguments specific to that software.

`set_args(object, ...)` - Modifies the arguments of a model specification

`set_mode(object, mode, …)` - Changes the model's mode.

`show_engines(x)` - The possible engines for a model can depend on what packages are loaded. Some parsnip extension add engines to existing models.

```{r}
show_engines("linear_reg")
```


