---
title: "Create models with parsnip :: Cheatsheet"
description: " "
image-alt: ""
execute:
  eval: true
  warning: false
---

```{css}
#| echo: false
.package-pills {
  display: flex;
  flex-wrap: wrap;
  gap: 8px;
  margin: 4px 0 10px 0;
}

.pill {
  display: inline-block;
  padding: 2px 12px;
  margin: 2px 0;
  border: 1px solid #6c757d;
  border-radius: 20px;
  font-size: 14px;
  font-family: monospace;
  color: #333;
  background-color: white;
  white-space: nowrap;
}

.pill sup {
  font-size: 10px;
  margin-left: 3px;
}
```

```{r}
#| output: asis
#| echo: false
#| column: margin
source("common.R")
use_cheatsheet_logo(
  "parsnip", 
  alt = "Hex logo for recipes, drawing of a cartoon parsnip smiling."
)
sheet_name <- tools::file_path_sans_ext(knitr::current_input())
pdf_preview_link(sheet_name)
translation_list(sheet_name)

library(tidymodels)

package_pills <- function(packages) {
  cat('::: {.package-pills}\n')
  for (pkg in packages) {
    name <- pkg[[1]]
    numbers <- pkg[[2]]
    sups <- paste0('<sup>', numbers, '</sup>', collapse = ' ')
    cat('<span class="pill">', name, ' ', sups, '</span>\n', sep = '')
  }
  cat(':::\n')
}
```

## Basics

`parsnip` provides a tidy, unified interface to a range of models from other packages. It helps avoid having to remember how to properly call the modeling functions of those external packages.

A `parsnip` specification is made up of 3 main components:

1.  The type of model to be used, such as Random Forest (`rand_forest()`) or linear regression (`linear_reg()`)

2.  How will the model be used, or mode. The two most common are “regression” and “classification”.

3.  The computational engine, or program that will actually execute the training. It could be an external R package, such as ranger, or even an engine outside of R, such as Stan or Apache Spark.

```{r}
library(tidymodels)

rand_forest(mtry = 10, trees = 2000) |> # Define type of model
  set_engine("ranger", importance = "impurity") |> # Select an engine
  set_mode("regression") # Set the mode
```

`set_engine(object, engine, ...)` - Specifies which package or system will be used to fit the model, along with any arguments specific to that software.

`set_args(object, ...)` - Modifies the arguments of a model specification

`set_mode(object, mode, …)` - Changes the model's mode.

`show_engines(x)` - The possible engines for a model can depend on what packages are loaded. Some parsnip extension add engines to existing models.

```{r}
show_engines("linear_reg")
```

## Legends

### Mode Support Numbers

- **1** - Classification
- **2** - Regression
- **3** - Censored Regression
- **4** - Quantile Regression

### Engine Tags

Engine tags show the engine name and mode support numbers. For example, <span class="pill">h2o <sup>12</sup></span> means engine `h2o` supports classification (1), and regression (2).



## Classification Only

-   `logistic_reg(mode = "classification", engine = "glm", penalty, mixture)` - Generalized linear model for binary outcomes. A linear combination of the predictors is used to model the log odds of an event.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("brulee", 1),
      list("gee", 1),
      list("glm", 1),
      list("glmer", 1),
      list("glmnet", 1),
      list("h2o", 1),
      list("keras", 1),
      list("LiblineaR", 1),
      list("spark", 1),
      list("stan", 1),
      list("stan_glmer", 1)
    ))
    ```
-   `multinom_reg(mode = "classification", engine = "nnet", penalty, mixture)` - Uses linear predictors to predict multiclass data using the multinomial distribution.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("brulee", 1),
      list("glmnet", 1),
      list("h2o", 1),
      list("keras", 1),
      list("nnet", 1),
      list("spark", 1)
    ))
    ```

-   `naive_Bayes(mode = "classification", smoothness, Laplace, engine = "klaR")` - Uses Bayes' theorem to compute the probability of each class, given the predictor values.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("h2o", 1),
      list("klaR", 1),
      list("naivebayes", 1)
    ))
    ```

-   `null_model(mode = "classification", engine = "parsnip")` - Fit a single mean or largest class model. This is the user-facing function for the null_model() specification.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("parsnip", 1)
    ))
    ```

-   `ordinal_reg(mode = "classification", ordinal_link, odds_link, penalty, mixture, engine = "polr")` - Defines a generalized linear model that predicts an ordinal outcome.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("rpartScore", 1),
      list("polr", 1),
      list("vgam", 1),
      list("vglm", 1)
    ))
    ```

## Regression Only

-   `linear_reg(mode = "regression", engine = "lm", penalty, mixture)` - Defines a model that can predict numeric values from predictors using a linear function.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("brulee", 2),
      list("gee", 2),
      list("glm", 2),
      list("glmer", 2),
      list("glmnet", 2),
      list("gls", 2),
      list("h2o", 2),
      list("keras", 2),
      list("lm", 2),
      list("lme", 2),
      list("quantreg", 2),
      list("spark", 2),
      list("stan", 2),
      list("stan_glmer", 2)
    ))
    ```

-   `poisson_reg(mode = "regression", penalty, mixture, engine = "glm")` - Defines a generalized linear model for count data that follow a Poisson distribution.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("gee", 2),
      list("glm", 2),
      list("glmer", 2),
      list("glmnet", 2),
      list("h2o", 2),
      list("hurdle", 2),
      list("stan", 2),
      list("stan_glmer", 2),
      list("zeroinfl", 2)
    ))
    ```

## General Use

-   `decision_tree(mode, engine = "rpart", cost_complexity, tree_depth, min_n)` - A set of if/then statements creates a tree-based structure.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("partykit", 123),
      list("rpart", 123),
      list("spark", 12),
      list("C5.0", 1)
    ))
    ```

-   `mars(mode, engine = "earth", num_terms, prod_degree, prune_method)` - Uses artificial features for some predictors. These features resemble hinge functions and the result is a model that is a segmented regression in small dimensions.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("earth", 12)
    ))
    ```

-   `mlp(mode, engine = "nnet", hidden_units, penalty, dropout, epochs, activation, learn_rate)` - Defines a multilayer perceptron model (a.k.a. a single layer, feed-forward neural network).

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("nnet", 12),
      list("brulee", 12),
      list("brulee_two_layer", 12),
      list("keras", 12),
      list("grnn", 12)
    ))
    ```

-   `gen_additive_mod(mode, select_features, adjust_deg_free, engine = "mgcv")` - Uses smoothed functions of numeric predictors in a generalized linear model.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("mgcv", 12)
    ))
    ```

-   `nearest_neighbor(mode, engine = "kknn", neighbors, weight_func, dist_power)` - Uses the K most similar data points from the training set to predict new samples.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("knn", 12)
    ))
    ```

-   `pls(mode, predictor_prop, num_comp, engine = "mixOmics")` - Uses latent variables to model the data. Similar to a supervised version of PCA.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("mixOmics", 12)
    ))
    ```

## Discriminant

-   `discrim_flexible(mode = "classification", num_terms, prod_degree, prune_method, engine = "earth")` - Fits a discriminant analysis model that uses nonlinear features created using MARS.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("earth", 1)
    ))
    ```

-   `discrim_regularized(mode = "classification", frac_common_cov, frac_identity, engine = "klaR")` - Estimates a multivariate distribution for the predictors separately for the data in each class. The model's structure can be LDA, QDA, or a combination. Each probability class is computed using Bayes's theorem, given the predictor values.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("klaR", 1)
    ))
    ```


<div style="text-align: center;">
Estimates a multivariate distribution for the predictors separately for the data in each class using a method described below. Each class' probability is computed using Bayes' theorem, given the predictor values.
</div>

-   `discrim_linear(mode = "classification", regularization_method, engine = "MASS", penalty)` - Uses Gaussian with a common covariance matrix to perform the estimate.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("MASS", 1),
      list("mda", 1),
      list("sda", 1),
      list("sparsediscrim", 1)
    ))
    ```

-   `discrim_quad(mode = "classification", regularization_method, engine = "MASS")` - Uses Gaussian with separate covariance matrices to perform the estimate.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("MASS", 1),
      list("sparsediscrim", 1)
    ))
    ```

## Support Vector Machine

<div style="text-align: center;">
**Classification:** Maximizes the width of the margin between classes using a method described below.

**Regression:** Optimizes a robust loss function only affected by very large model residuals and uses an additional method described below.
</div>

-   `svm_linear(mode, cost, engine = "LiblineaR", margin)` - Classification: A linear class boundary. Regression: Uses a linear fit.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("kernlab", 12),
      list("LiblineaR", 12)
    ))
    ```

-   `svm_poly(mode, cost, engine = "kernlab", degree, scale_factor)` - Classification: A polynomial class boundary. Regression: Uses polynomial functions of the predictors.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("kernlab", 12)
    ))
    ```

-   `svm_rbf(mode, cost, engine = "kernlab", rbf_sigma)` - Classification: A nonlinear class boundary. Regression: Uses nonlinear functions of the predictors.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("kernlab", 12)
    ))
    ```

## Feature Rules

-   `rule_fit(mode, mtry, trees, min_n, tree_depth, learn_rate, loss_reduction, sample_size, stop_iter, penalty, engine = "xrf")` - Derives simple feature rules from a tree ensemble and uses them as features in a regularized model.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("xrf", 12),
      list("h2o", 1)
    ))
    ```

-   `C5_rules(mode = "classification", trees, min_n, engine = "C5.0")` - Derives feature rules from a tree for prediction. A single tree or boosted ensemble can be used.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("C5.0", 1)
    ))
    ```

-   `cubist_rules(mode = "regression", committees, neighbors, max_rules, engine = "Cubist")` - Derives simple feature rules from a tree ensemble and creates regression models within each rule.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("Cubist", 2)
    ))
    ```

## Ensemble

<div style="text-align: center; font-style: italic;">
"E Pluribus Unum"
</div>

-   `bag_mars(mode, num_terms, prod_degree, prune_method, engine = "earth")` - Ensemble of generalized linear models that use artificial features for some predictors. These features resemble hinge functions and the result is a model that is a segmented regression in small dimensions.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("earth", 12)
    ))
    ```

-   `bag_mlp(mode, hidden_units, penalty, epochs, engine = "nnet")` - An ensemble of single layer, feed-forward neural networks.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("nnet", 12)
    ))
    ```

-   `bag_tree(mode, cost_complexity = 0, tree_depth, min_n = 2, class_cost, engine = "rpart")` - Ensemble of decision trees.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("C5.0", 1),
      list("rpart", 123)
    ))
    ```

-   `bart(mode, engine = "dbarts", trees, prior_terminal_node_coef, prior_terminal_node_expo, prior_outcome_range)` - Tree ensemble model that uses Bayesian analysis to assemble the ensemble.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("bart", 12)
    ))
    ```

-   `boost_tree(mode, engine = "xgboost", mtry, trees, min_n, tree_depth, learn_rate, loss_reduction, sample_size, stop_iter)` - Creates a series of decision trees forming an ensemble. Each tree depends on the results of previous trees. All trees in the ensemble are combined to produce a final prediction.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("C5.0", 1),
      list("catboost", 12),
      list("h2o", 12),
      list("lightgbm", 12),
      list("mboost", 3),
      list("spark", 12),
      list("xgboost", 124)
    ))
    ```

-   `rand_forest(mode, engine = "ranger", mtry, trees, min_n)` - Creates a large number of decision trees, each independent of the others. The final prediction uses all predictions from the individual trees and combines them.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("aorsf", 123),
      list("grf", 124),
      list("h2o", 12),
      list("partykit", 123),
      list("randomForest", 12),
      list("ranger", 12),
      list("spark", 12)
    ))
    ```

## Survival

**`*(mode = "censored regression")`**

-   `proportional_hazards(engine = "survival", penalty, mixture)` - Defines a model for the hazard function as a multiplicative function of covariates times a baseline hazard.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("glmnet", 3),
      list("survival", 3)
    ))
    ```

-   `survival_reg(engine = "survival", dist)` - Defines a parametric survival model.

    ```{r}
    #| output: asis
    #| echo: false
    package_pills(list(
      list("flexsurv", 3),
      list("flexsurvspline", 3),
      list("survival", 3)
    ))
    ```

## Operations

```{r}
library(tidymodels)
lm_spec <- linear_reg() |> set_engine("lm")
```

### Methods

-   `fit(object, ...)` - Estimates parameters for a given model from a set of data.

    ```{r}
    lm_fit <- fit(lm_spec, mpg ~ ., data = mtcars)
    ```

-   `predict(object, ...)`

    ```{r}
    predict(lm_fit, mtcars)
    ```

-   `autoplot(object, ...)` - Uses ggplot2 to draw a particular plot for an object of a particular class

-   `update(object, ...)` - Updates and (by default) re-fit a model. It does this by extracting the call stored in the object, updating the call and evaluating that call.

### Tidiers

-   `augment(x, ...)` - Augment data with model results

    ```{r}
    augment(lm_fit, mtcars)
    ```

-   `glance(x, ...)` - Construct a single row summary "glance" of a model fit

    ```{r}
    glance(lm_fit)
    ```

-   `tidy(x, ...)` - Turn an object into a tidy tibble

    ```{r}
    tidy(lm_fit)
    ```

### General

-   `repair_call(x, data)` - When the user passes a formula to fit() and the underlying model function uses a formula, the call object produced by fit() may not be usable by other functions.

-   `control_parsnip(verbosity = 1L, catch = FALSE)` - Pass options to the fit.model_spec() function to control its output and computations.

    ```{r}
    control_parsnip(verbosity = 2)
    ```

-   `show_engines(x)` - The possible engines for a model can depend on what packages are loaded. Some parsnip extension add engines to existing models.

    ```{r}
    show_engines("linear_reg")
    ```

-   `translate(x, ...)` - Translates a model specification into a code object that is specific to a particular engine (e.g. R package). It translates generic parameters to their counterparts.

    ```{r}
    translate(lm_spec)
    ```

-   `multi_predict(object, ...)` - For some models, predictions can be made on sub-models in the model object.

### Extract

-   `extract_spec_parsnip(x, ...)` - Returns a parsnip model specification.

    ```{r}
    extract_spec_parsnip(lm_fit)
    ```

-   `extract_fit_engine(x, ...)` - Returns the engine specific fit embedded within a parsnip model fit. For example, when using linear_reg() with the "lm" engine, this returns the underlying lm object.

    ```{r}
    extract_fit_engine(lm_fit)
    ```

-   `extract_parameter_dials(x, parameter, ...)` - Returns a single dials parameter object.

-   `extract_parameter_set_dials(x, ...)` - Returns a set of dials parameter objects.

-   `extract_fit_time(x, summarize = TRUE, ...)` - returns a tibble with fit times. The fit times correspond to the time for the parsnip engine to fit and do not include other portions of the elapsed time in fit.model_spec().
