---
title: "Preprocessing data with recipes :: Cheatsheet"
description: " "
image-alt: ""
execute:
  eval: true
  output: false
  warning: false
---

```{r}
#| output: asis
#| echo: false
#| column: margin
source("common.R")
use_cheatsheet_logo(
  "recipes", 
  alt = "Hex logo for recipes, drawing of a cartoon cupcake that has arms and legs and it is smiling ."
)
sheet_name <- tools::file_path_sans_ext(knitr::current_input())
pdf_preview_link(sheet_name)
translation_list(sheet_name)
```

## Basics

Get your data ready for modeling using ‘pipable’ sequences of feature engineering steps with **recipes.**

```{r}
#| eval: false

# Initialize the recipe and add steps 
rec <- recipe(x ~ ., data = train_data) |>
  step_normalize(all_numeric_predictors())

# Run the steps using training data
pr <- prep(rec, training = train_data)

#  Apply estimates to new data 
bake(pr, new_data = new_data)
```

-   `recipe(x, ...)`: Begins a new recipe specification.

-   `prep(x, ...)`: Prepares the recipe with training data.

-   `bake(object, ...)`: Applies estimates from prep().

-   `update(object, ...)`: Updates and re-fits a model.

### Common `step_` arguments

|  |  |
|----------|-----------------------------------------------------------|
| `recipe` | A recipe object. New steps are appended to the recipe. |
| `...` | Arguments passed to the external R function accessed by the step function |
| `options` | Selector functions to choose variables for this step |

## Filters

-   `step_nzv(recipe, ..., freq_cut = 95/5, unique_cut = 10, options = list(freq_cut = 95/5)`: Removes variables that are highly sparse and unbalanced.

-   `step_zv(recipe, ..., group = NULL)`: Removes variables that contain only a single value.

-   `step_lincomb(recipe, ..., max_steps = 5)`: Removes numeric variables that have exact linear combinations between them.

-   `step_corr(recipe, ..., threshold = 0.9, use = "pairwise.complete.obs", method = "pearson")`: Removes variables that have large absolute correlations with other variables.

-   `step_filter_missing(recipe, ..., threshold = 0.1)`: Removes variables that have too many missing values.

-   `step_rm(recipe, ...)`: Removes selected variables.

## In-place Transformations

- `step_mutate(recipe, ..., .pkgs = character())`: General purpose transformer using `dplyr.`

- `step_relu(recipe, ..., shift = 0, reverse = FALSE, smooth = FALSE, prefix = "right_relu_")`: Applies smoothed rectified linear transformation.

- `step_sqrt(recipe, ...)`: Applies square root transformation.

### Basis functions

- `step_spline_natural(recipe, ..., deg_free = 10, options = NULL, keep_original_cols = FALSE)`: Creates a natural spline (a.k.a restricted cubic spline) features.

- `step_spline_b(recipe, ..., deg_free = 10, degree = 3, options = NULL, keep_original_cols = FALSE)`: Creates b-spline features.

- `step_spline_convex(recipe, ..., deg_free = 10, degree = 3,  options = NULL, keep_original_cols = FALSE)`

- `step_spline_monotone(recipe, ..., deg_free = 10, degree = 3, options = NULL, keep_original_cols = FALSE)` 

- `step_spline_nonnegative(recipe, ..., deg_free = 10, degree = 3, options = NULL, keep_original_cols = FALSE)` 

- `step_poly(recipe, ..., degree = 2L, options = list(), keep_original_cols = FALSE)`: Creates new columns that are basis expansions of variables using orthogonal polynomials.

- `step_poly_bernstein(recipe, ..., degree = 10, options = NULL, results = NULL, keep_original_cols = FALSE)`: Creates Bernstein polynomial features.

### Normalization

- `step_normalize(recipe, ..., na_rm = TRUE)`: Normalizes to have a standard deviation of 1 and mean of 0.

- `step_YeoJohnson(recipe, ...)`: Makes data look more like a normal distribution.

- `step_percentile(recipe, ..., options = list(probs = (0:100)/100), outside = "none")`: Replaces the value of a variable with its percentile from the training set.

- `step_range(recipe, ..., min = 0, max = 1, clipping = TRUE)`: Normalizes numeric data to be within a pre-defined range of values.

- `step_spatialsign(recipe, ..., na_rm = TRUE)`: Converts numeric data into a projection on to a unit sphere.

### Discretize

- `step_discretize(recipe, ..., num_breaks = 4, min_unique = 10, options = list(prefix = "bin"))`: Converts numeric data into a factor with bins having approximately the same number of data points.

- `step_cut(recipe, ..., breaks, include_outside_range = FALSE)`: Cuts a numeric variable into a factor based on provided boundary values.

