---
title: "Preprocessing data with recipes :: Cheatsheet"
description: " "
image-alt: ""
execute:
  eval: true
  output: false
  warning: false
---

```{r}
#| output: asis
#| echo: false
#| column: margin
source("common.R")
use_cheatsheet_logo(
  "recipes", 
  alt = "Hex logo for recipes, drawing of a cartoon cupcake that has arms and legs and it is smiling ."
)
sheet_name <- tools::file_path_sans_ext(knitr::current_input())
pdf_preview_link(sheet_name)
translation_list(sheet_name)
```

## Basics

Get your data ready for modeling using ‘pipable’ sequences of feature engineering steps with **recipes.**

```{r}
#| eval: false

# Initialize the recipe and add steps 
rec <- recipe(x ~ ., data = train_data) |>
  step_normalize(all_numeric_predictors())

# Run the steps using training data
pr <- prep(rec, training = train_data)

#  Apply estimates to new data 
bake(pr, new_data = new_data)
```

-   `recipe(x, ...)`: Begins a new recipe specification.

-   `prep(x, ...)`: Prepares the recipe with training data.

-   `bake(object, ...)`: Applies estimates from prep().

-   `update(object, ...)`: Updates and re-fits a model.

### Common `step_` arguments

|  |  |
|---------------|---------------------------------------------------------|
| `recipe` | A recipe object. New steps are appended to the recipe. |
| `...` | Arguments passed to the external R function accessed by the step function |
| `options` | Selector functions to choose variables for this step |

## Filters

-   `step_nzv(recipe, ..., freq_cut = 95/5, unique_cut = 10, options = list(freq_cut = 95/5)`: Removes variables that are highly sparse and unbalanced.

-   `step_zv(recipe, ..., group = NULL)`: Removes variables that contain only a single value.

-   `step_lincomb(recipe, ..., max_steps = 5)`: Removes numeric variables that have exact linear combinations between them.

-   `step_corr(recipe, ..., threshold = 0.9, use = "pairwise.complete.obs", method = "pearson")`: Removes variables that have large absolute correlations with other variables.

-   `step_filter_missing(recipe, ..., threshold = 0.1)`: Removes variables that have too many missing values.

-   `step_rm(recipe, ...)`: Removes selected variables.

## In-place Transformations

-   `step_mutate(recipe, ..., .pkgs = character())`: General purpose transformer using `dplyr.`

-   `step_relu(recipe, ..., shift = 0, reverse = FALSE, smooth = FALSE, prefix = "right_relu_")`: Applies smoothed rectified linear transformation.

-   `step_sqrt(recipe, ...)`: Applies square root transformation.

### Basis functions

-   `step_spline_natural(recipe, ..., deg_free = 10, options = NULL, keep_original_cols = FALSE)`: Creates a natural spline (a.k.a restricted cubic spline) features.

-   `step_spline_b(recipe, ..., deg_free = 10, degree = 3, options = NULL, keep_original_cols = FALSE)`: Creates b-spline features.

-   `step_spline_convex(recipe, ..., deg_free = 10, degree = 3,  options = NULL, keep_original_cols = FALSE)`

-   `step_spline_monotone(recipe, ..., deg_free = 10, degree = 3, options = NULL, keep_original_cols = FALSE)`

-   `step_spline_nonnegative(recipe, ..., deg_free = 10, degree = 3, options = NULL, keep_original_cols = FALSE)`

-   `step_poly(recipe, ..., degree = 2L, options = list(), keep_original_cols = FALSE)`: Creates new columns that are basis expansions of variables using orthogonal polynomials.

-   `step_poly_bernstein(recipe, ..., degree = 10, options = NULL, results = NULL, keep_original_cols = FALSE)`: Creates Bernstein polynomial features.

### Normalization

-   `step_normalize(recipe, ..., na_rm = TRUE)`: Normalizes to have a standard deviation of 1 and mean of 0.

-   `step_YeoJohnson(recipe, ...)`: Makes data look more like a normal distribution.

-   `step_percentile(recipe, ..., options = list(probs = (0:100)/100), outside = "none")`: Replaces the value of a variable with its percentile from the training set.

-   `step_range(recipe, ..., min = 0, max = 1, clipping = TRUE)`: Normalizes numeric data to be within a pre-defined range of values.

-   `step_spatialsign(recipe, ..., na_rm = TRUE)`: Converts numeric data into a projection on to a unit sphere.

### Discretize

-   `step_discretize(recipe, ..., num_breaks = 4, min_unique = 10, options = list(prefix = "bin"))`: Converts numeric data into a factor with bins having approximately the same number of data points.

-   `step_cut(recipe, ..., breaks, include_outside_range = FALSE)`: Cuts a numeric variable into a factor based on provided boundary values.


## Imputation

-   `step_impute_bag(recipe, ..., impute_with = all_predictors(), trees = 25, options = list(keepX = FALSE))`:  Creates a bagged tree model for data. Good for categorical data.

-   `step_impute_knn(recipe, ..., neighbors = 5, impute_with = all_predictors(), options = list(nthread = 1, eps = 1e-08))`:  Uses Gower's distance which can be used for mixtures of nominal and numeric data.

-   `step_impute_linear(recipe, ..., impute_with = all_predictors())`:  Creates linear regression models to impute missing data.

-   `step_impute_lower(recipe, ..., threshold = NULL)`:  Substitutes the truncated value by a random number between zero and the truncation point.

-   `step_impute_mean(recipe, ..., trim = 0)`:  Substitutes missing values of numeric variables by the training set mean of those variables.

-   `step_impute_median(recipe, ...)`:  Substitutes missing values of numeric variables by the training set median of those variables.

- `step_impute_mode(recipe, ...)`:  Imputes nominal data using the most common value.

- `step_impute_roll(recipe, ..., statistic = median, window = 5L)`:  Imputes numeric data using a rolling window statistic.

- `step_unknown(recipe, ..., new_level = "unknown")`:  Assigns a missing value in a factor level to “unknown”.


## Encodings

### Type Converters

- `step_factor2string(recipe, ...)`: Converts one or more factor vectors to strings.

- `step_string2factor(recipe, ...)`: Converts one or more character vectors to factors (ordered or unordered).

- `step_num2factor(recipe, ..., transform = function(x) x)`: Converts one or more numeric vectors to factors (ordered or unordered). This can be useful when categories are encoded as integers.

- `step_integer(recipe, ..., strict = TRUE, zero_based = FALSE)`: Converts data into a set of ascending integers based on the ascending order from the training data.

### Value Converters


- `step_indicate_na(recipe, ..., sparse = "auto", keep_original_cols = TRUE)`: Creates and append additional binary columns to the data set to indicate which observations are missing.

- `step_ordinalscore(recipe, ..., convert = as.numeric)`: Converts ordinal factor variables into numeric scores.

- `step_unorder(recipe, ...)`: Turns ordered factor variables into unordered factor variables.


### Other

- `step_relevel(recipe, ..., ref_level)`: Reorders factor columns so that the level specified by ref_level is first. This is useful for `contr.treatment()` contrasts which take the first level as the reference.

- `step_novel(recipe, ..., new_level = "new")`: Assigns a previously unseen factor level to "new" .

- `step_other(recipe, ..., threshold = 0.05, other = "other" )`: Pools infrequently occurring values into an "other" category.


