---
title: "Preprocessing data with recipes :: Cheatsheet"
description: " "
image-alt: ""
execute:
  eval: true
  output: false
  warning: false
---

```{r}
#| output: asis
#| echo: false
#| column: margin
source("common.R")
use_cheatsheet_logo(
  "recipes", 
  alt = "Hex logo for recipes, drawing of a cartoon cupcake that has arms and legs and it is smiling ."
)
sheet_name <- tools::file_path_sans_ext(knitr::current_input())
pdf_preview_link(sheet_name)
translation_list(sheet_name)
```

## Basics

Get your data ready for modeling using ‘pipable’ sequences of feature engineering steps with **recipes.**

```{r}
#| eval: false

# Initialize the recipe and add steps 
rec <- recipe(x ~ ., data = train_data) |>
  step_normalize(all_numeric_predictors())

# Run the steps using training data
pr <- prep(rec, training = train_data)

#  Apply estimates to new data 
bake(pr, new_data = new_data)
```

-   `recipe(x, ...)`: Begins a new recipe specification.

-   `prep(x, ...)`: Prepares the recipe with training data.

-   `bake(object, ...)`: Applies estimates from prep().

-   `update(object, ...)`: Updates and re-fits a model.

### Common `step_` arguments

|  |  |
|----------|-----------------------------------------------------------|
| `recipe` | A recipe object. New steps are appended to the recipe. |
| `...` | Arguments passed to the external R function accessed by the step function |
| `options` | Selector functions to choose variables for this step |

## Filters

-   `step_nzv(recipe, ..., freq_cut = 95/5, unique_cut = 10, options = list(freq_cut = 95/5)`: Removes variables that are highly sparse and unbalanced.

-   `step_zv(recipe, ..., group = NULL)`: Removes variables that contain only a single value.

-   `step_lincomb(recipe, ..., max_steps = 5)`: Removes numeric variables that have exact linear combinations between them.

-   `step_corr(recipe, ..., threshold = 0.9, use = "pairwise.complete.obs", method = "pearson")`: Removes variables that have large absolute correlations with other variables.

-   `step_filter_missing(recipe, ..., threshold = 0.1)`: Removes variables that have too many missing values.

-   `step_rm(recipe, ...)`: Removes selected variables.
