---
title: "Reproducible R Environments with renv :: Cheatsheet"
description: " "
image-alt: ""
execute:
  eval: true
  output: false
  warning: false
---

```{r}
#| output: asis
#| echo: false
#| column: margin
source("common.R")
use_cheatsheet_logo(
  "renv",
  alt = "Hex logo for renv with a leafy plant growing out of the soil and the roots visible below the soil."
)
sheet_name <- tools::file_path_sans_ext(knitr::current_input())
pdf_preview_link(sheet_name)
translation_list(sheet_name)
```

The **renv** package helps you create isolated, portable, and reproducible environments for your R projects.
Use as little or as much of renv as you need -- if all you want is project isolation, then `init()` is all you need. 
All of the "regular" tools for R package management and installation continue to work with renv.
renv works with typical R package repositories (CRAN, Posit Package Manager, R-universe) and common Git services (GitHub, GitLab, Bitbucket). 

## Core Concepts

- An R **library** is a directory of installed R packages. When you install a package (e.g. via `install.packages()`), it goes into the first writable library on your `.libPaths()`.

- renv projects use an *isolated* **project library**. Installing or removing packages in one renv project does not affect other renv projects, or your default R library.

- The **lockfile** (`renv.lock`) records the exact source and version of every package your project uses. `snapshot()` writes it; `restore()` reads it.

- The global **package cache** is shared across all of your renv projects. Project libraries are populated with links into the cache, so creating a new project library is cheap, and installing a package you've used before is fast.

## The renv Workflow

The core reproducibility loop ties every function together:

1.  `init()` -- set up an isolated project library.
2.  `install()` / `update()` -- add and update packages as you work.
3.  Confirm your code still works, then `snapshot()` -- record the exact package versions to the **lockfile** (`renv.lock`).
4.  Commit the project files to version control and share it (see **Collaboration** below).
5.  A collaborator (or future you) runs `restore()` to reinstall those exact versions.

## Collaboration

Commit these files to version control so collaborators can reproduce your project:

- `renv.lock` -- the lockfile of exact package versions.
- `.Rprofile` -- activates renv for new R sessions in the project.
- `renv/activate.R` -- the activation script.
- `renv/settings.json` -- project-level settings.

renv writes a `.gitignore` for you so the project library itself is not committed -- you share the lockfile, not the installed packages. When a collaborator opens the project, renv automatically prompts them to run `restore()`.

The same workflow powers CI: [`r-lib/actions/setup-renv`](https://github.com/r-lib/actions/tree/v2-branch/setup-renv) restores your lockfile on GitHub Actions; see the [CI vignette](https://rstudio.github.io/renv/articles/ci.html) for other services.

## Initialize

Call `init()` on a new or existing project to initialize renv.

- `init()` creates a project `.Rprofile` and `renv/activate.R`, which automatically activate renv for new R sessions started in the project.

- Set project options at init time, e.g. to use explicit snapshots:

```{r}
#| eval: false
init(settings = list(snapshot.type = "explicit"))
```

## Install & Update Packages

Use these to change *which* packages (and versions) are in your project library:

- `install()`: Install one or more named packages. Supports custom remotes, e.g. `install("tidyverse/dplyr")` (GitHub) or a specific version with `install("dplyr@1.1.0")`. With no arguments, installs the packages your project requires.
- `update()`: Update packages (all by default, or those you name) to the latest version available from the source they were installed from.
- `remove()`: Remove packages from the project library.

Enable [pak](https://pak.r-lib.org) for faster, parallel installs with `options(renv.config.pak.enabled = TRUE)`.

## Snapshot & Restore

Use these to save and reproduce the *state* of your project library:

- `status()`: Check whether your lockfile and library are in sync (see **Staying in Sync**).
- `snapshot()`: Save the state of your project library to the lockfile.
- `restore()`: Install the exact package versions recorded in the lockfile.
- `record()`: Record a specific package version in the lockfile without installing it.

## Staying in Sync

`status()` reports when your project is out of sync, and how to fix it:

- Packages installed or updated, but not recorded in the lockfile? Record them with `snapshot()`.
- Packages recorded in the lockfile, but not installed? Reinstall them with `restore()`.
- Packages used in your code, but not installed? Install them with `install()`, then `snapshot()`.
- Pulled a collaborator's changes to `renv.lock`? Sync up with `restore()`.

When in doubt, treat the lockfile as the source of truth -- your project library can always be rebuilt from it with `restore()`.

## Dependencies

`dependencies()`: Discover the packages used in your project's R, R Markdown, and Quarto files.
The rmarkdown package is required to scan dependencies in R Markdown and Quarto files.

Use a `.renvignore` file to exclude files or directories from dependency discovery, and `settings$ignored.packages()` to keep development-only packages (e.g. devtools) out of the lockfile.

renv supports several **snapshot types**, set with `settings$snapshot.type()`. The two most common:

| Type | What it captures |
|-------------------|-----------------------------------------------------|
| `"implicit"` (default) | Scans your files with `dependencies()` (e.g. `library(dplyr)`, `dplyr::mutate()`). |
| `"explicit"` | Only packages listed in the project `DESCRIPTION` file, which you maintain. |

## Configuring renv

- `?config`: User-specific settings. Tweak how renv behaves in different scenarios.
- `?settings`: Project-level settings. Persisted with the project and active for all collaborators.
- `deactivate()` / `activate()`: Temporarily turn renv off in a project, and turn it back on.

## Other Useful Functions

- `history()` / `revert()`: View past lockfiles from version control, and roll back to an earlier one.
- `upgrade()`: Upgrade the version of renv used by the project.
- `repair()`: Fix a project library whose links into the global cache have broken (e.g. after moving or cleaning the cache).
- `purge()`: Remove packages from the global cache (e.g. to reclaim space or clear a broken install).
- `diagnostics()`: Print a diagnostics report; useful for bug reports at [github.com/rstudio/renv](https://github.com/rstudio/renv).

## What renv Does Not Do

renv reproduces R packages, not your whole environment. It does not manage:

- Your R version -- renv records it in the lockfile and warns on mismatch, but does not install it. Use a tool like [rig](https://github.com/r-lib/rig) to install and switch R versions.
- Pandoc -- required to render R Markdown and Quarto documents.
- System libraries, the OS, and compiler versions -- for full reproducibility, pair renv with Docker.

`restore()` can also fail when a package must be compiled from source and its system prerequisites (or original binaries) are unavailable.

------------------------------------------------------------------------

CC BY SA Posit Software, PBC • [info\@posit.co](mailto:info@posit.co) • [posit.co](https://posit.co)

Learn more at [rstudio.github.io/renv](https://rstudio.github.io/renv/)

Updated: `r format(Sys.Date(), "%Y-%m")`.

```{r}
#| output: true
packageVersion("renv")
```

------------------------------------------------------------------------
