---
title: "Reproducible R Environments with renv :: Cheatsheet"
description: " "
image-alt: ""
execute:
  eval: true
  output: false
  warning: false
---

```{r}
#| output: asis
#| echo: false
#| column: margin
source("common.R")
use_cheatsheet_logo(
  "renv",
  alt = "Hex logo for renv with a leafy plant growing out of the soil and the roots visible below the soil."
)
sheet_name <- tools::file_path_sans_ext(knitr::current_input())
# TODO: re-enable once the PDF version (renv.pdf, pngs/renv.png) exists
# pdf_preview_link(sheet_name)
translation_list(sheet_name)
```

<!--
Deferred to the pkgdown website (https://rstudio.github.io/renv), likely too much for the sheet:
- renv infers a package's source from its DESCRIPTION file, so the installer (renv, pak, remotes) must encode Remote dependency info at install time.
- renv::run() (run an R script in the context of a project).
- renv::embed() / renv::use() (declare dependency versions inline in .R / .Rmd / .qmd files).

Still to confirm / possible additions:
- renv::checkout() (install packages from a snapshot of Posit Package Manager).
-->

The **renv** package helps you create isolated, portable, and reproducible environments for your R projects.
Use as little or as much of renv as you need -- if all you want is project isolation, then `renv::init()` is all you need. 
All of the "regular" tools for R package management and installation continue to work with renv.
renv works with typical R package repositories (CRAN, Posit Package Manager, R-universe) and common Git services (GitHub, GitLab, Bitbucket). 

## Core Concepts

- An R **library** is a directory of installed R packages. When you install a package (e.g. via `install.packages()`), it goes into the first writable library on your `.libPaths()`.

- renv projects use an *isolated* **project library**. Installing or removing packages in one renv project does not affect other renv projects, or your default R library.

- The **lockfile** (`renv.lock`) records the exact source and version of every package your project uses. `renv::snapshot()` writes it; `renv::restore()` reads it.

- The global **package cache** is shared across all of your renv projects. Project libraries are populated with links into the cache, so creating a new project library is cheap, and installing a package you've used before is fast.

## The renv Workflow

The core reproducibility loop ties every function together:

1.  `renv::init()` -- set up an isolated project library.
2.  `renv::install()` / `renv::update()` -- add and update packages as you work.
3.  Confirm your code still works, then `renv::snapshot()` -- record the exact package versions to the **lockfile** (`renv.lock`).
4.  Commit the project files to version control and share it (see **Collaboration** below).
5.  A collaborator (or future you) runs `renv::restore()` to reinstall those exact versions.

## Collaboration

Commit these files to version control so collaborators can reproduce your project:

- `renv.lock` -- the lockfile of exact package versions.
- `.Rprofile` -- activates renv for new R sessions in the project.
- `renv/activate.R` -- the activation script.
- `renv/settings.json` -- project-level settings.

renv writes a `.gitignore` for you so the project library itself is not committed -- you share the lockfile, not the installed packages. When a collaborator opens the project, renv automatically prompts them to run `renv::restore()`.

The same workflow powers CI: [`r-lib/actions/setup-renv`](https://github.com/r-lib/actions/tree/v2-branch/setup-renv) restores your lockfile on GitHub Actions; see the [CI vignette](https://rstudio.github.io/renv/articles/ci.html) for other services.

## Getting Started

- Call `renv::init()` on a new or existing project to initialize renv.

- `renv::init()` creates a project `.Rprofile` and `renv/activate.R`, which automatically activate renv for new R sessions started in the project.

- Set project options at init time, e.g. to use explicit snapshots:

```{r}
#| eval: false
renv::init(settings = list(snapshot.type = "explicit"))
```

## Install & Update Packages

Use these to change *which* packages (and versions) are in your project library:

- `renv::install()`: Install one or more named packages. Supports custom remotes, e.g. `renv::install("tidyverse/dplyr")` (GitHub) or a specific version with `renv::install("dplyr@1.1.0")`. With no arguments, installs the packages your project requires.
- `renv::update()`: Update packages (all by default, or those you name) to the latest version available from the source they were installed from.
- `renv::remove()`: Remove packages from the project library.

Enable [pak](https://pak.r-lib.org) for faster, parallel installs with `options(renv.config.pak.enabled = TRUE)`.

## Snapshot & Restore

Use these to save and reproduce the *state* of your project library:

- `renv::status()`: Check whether your lockfile and library are in sync. Packages installed but not recorded call for `snapshot()`; packages recorded but not installed call for `restore()`.
- `renv::snapshot()`: Save the state of your project library to the lockfile.
- `renv::restore()`: Install the exact package versions recorded in the lockfile.
- `renv::record()`: Record a specific package version in the lockfile without installing it.

## Dependencies

`renv::dependencies()`: Discover the packages used in your project's R, R Markdown, and Quarto files.
The rmarkdown package is required to scan dependencies in R Markdown and Quarto files.

Use a `.renvignore` file to exclude files or directories from dependency discovery, and `renv::settings$ignored.packages()` to keep development-only packages (e.g. devtools) out of the lockfile.

renv supports several **snapshot types**, set with `renv::settings$snapshot.type()`. The two most common:

| Type | What it captures |
|-------------------|-----------------------------------------------------|
| `"implicit"` (default) | Scans your files with `renv::dependencies()` (e.g. `library(dplyr)`, `dplyr::mutate()`). |
| `"explicit"` | Only packages listed in the project `DESCRIPTION` file, which you maintain. |

## Configuring renv

- `?renv::config`: User-specific settings. Tweak how renv behaves in different scenarios.
- `?renv::settings`: Project-level settings. Persisted with the project and active for all collaborators.
- `renv::deactivate()` / `renv::activate()`: Temporarily turn renv off in a project, and turn it back on.

## Other Useful Functions

- `renv::history()` / `renv::revert()`: View past lockfiles from version control, and roll back to an earlier one.
- `renv::upgrade()`: Upgrade the version of renv used by the project.
- `renv::repair()`: Fix a project library whose links into the global cache have broken (e.g. after moving or cleaning the cache).
- `renv::purge()`: Remove packages from the global cache (e.g. to reclaim space or clear a broken install).
- `renv::diagnostics()`: Print a diagnostics report; useful for bug reports at [github.com/rstudio/renv](https://github.com/rstudio/renv).

## What renv Does Not Do

renv reproduces R packages, not your whole environment. It does not manage:

- Your R version -- renv records it in the lockfile and warns on mismatch, but does not install it. Use a tool like [rig](https://github.com/r-lib/rig) to install and switch R versions.
- Pandoc -- required to render R Markdown and Quarto documents.
- System libraries, the OS, and compiler versions -- for full reproducibility, pair renv with Docker.

`renv::restore()` can also fail when a package must be compiled from source and its system prerequisites (or original binaries) are unavailable.

------------------------------------------------------------------------

CC BY SA Posit Software, PBC • [info\@posit.co](mailto:info@posit.co) • [posit.co](https://posit.co)

Learn more at [rstudio.github.io/renv](https://rstudio.github.io/renv/)

Updated: `r format(Sys.Date(), "%Y-%m")`.

```{r}
#| output: true
packageVersion("renv")
```

------------------------------------------------------------------------
