---
title: "String manipulation with stringr :: Cheatsheet"
description: " "
image-alt: ""
execute:
  eval: true
  output: false
  warning: false
---

```{r}
#| output: asis
#| echo: false
#| column: margin
source("common.R")
use_cheatsheet_logo(
  "stringr", 
  alt = "Hex logo for stringr - a green hexagon with a white violin in the centre, above the word 'stringr' in white.")
sheet_name <- tools::file_path_sans_ext(knitr::current_input())
pdf_preview_link(sheet_name)
translation_list(sheet_name)
```

<!-- Page 1 -->

The **stringr** package provides a set of internally consistent tools for working with character strings, i.e. sequences of characters surrounded by quotation marks.

```{r}
library(stringr)
```

```{r}
#| include: false
see <- function(rx) str_view("abc ABC 123\t.!?\\(){}\n", rx)
```

## Detect Matches

-   `str_detect(string, pattern, negate = FALSE)`: Detect the presence of a pattern match in a string.
    Also `str_like()`.

    ```{r}
    str_detect(fruit, "a")
    ```

-   `str_starts(string, pattern, negate = FALSE)`: Detect the presence of a pattern match at the beginning of a string.
    Also `str_ends()`.

    ```{r}
    str_starts(fruit, "a")
    ```

-   `str_which(string, pattern, negate = FALSE)`: Find the indexes of strings that contain a pattern match.

    ```{r}
    str_which(fruit, "a")
    ```

-   `str_locate(string, pattern)`: Locate the positions of pattern matches in a string.
    Also `str_locate_all()`.

    ```{r}
    str_locate(fruit, "a")
    ```

-   `str_count(string, pattern)`: Count the number of matches in a string.

    ```{r}
    str_count(fruit, "a")
    ```

## Mutate Strings

-   `str_sub() <- value`: Replace substrings by identifying the substrings with `str_sub()` and assigning into the results.

    ```{r}
    str_sub(fruit, 1, 3) <- "str"
    ```

-   `str_replace(string, pattern, replacement)`: Replace the first matched pattern in each string.
    Also `str_remove()`.

    ```{r}
    str_replace(fruit, "p", "-")
    ```

-   `str_replace_all(string, pattern, replacement)`: Replace all matched patterns in each string.
    Also `str_remove_all()`.

    ```{r}
    str_replace_all(fruit, "p", "-")
    ```

-   `str_to_lower(string, locale = "en")`^1^: Convert strings to lower case.

    ```{r}
    str_to_lower(sentences)
    ```

-   `str_to_upper(string, locale = "en")`^1^: Convert strings to upper case.

    ```{r}
    str_to_upper(sentences)
    ```

-   `str_to_title(string, locale = "en")`^1^: Convert strings to title case.
    Also `str_to_setence()`.

    ```{r}
    str_to_title(sentences)
    ```

## Subset Strings

-   `str_sub(string, start = 1L, end = -1L)`: Extract substrings from a character vector.

    ```{r}
    str_sub(fruit, 1, 3)
    str_sub(fruit, -2)
    ```

-   `str_subset(string, pattern, negate = FALSE)`: Return only the strings that contain a pattern match.

    ```{r}
    str_subset(fruit, "p")
    ```

-   `str_extract(string, pattern)`: Return the first pattern match found in each string, as a vector.
    Also `str_extract_all()` to return every pattern match.

    ```{r}
    str_extract(fruit, "[aeiou]")
    ```

-   `str_match(string, pattern)`: Return the first pattern match found in each string, as a matrix with a column for each ( ) group in pattern.
    Also `str_match_all()`.

    ```{r}
    str_match(sentences, "(a|the) ([^ +])")
    ```

## Join and Split

-   `str_c(..., sep = "", collapse = NULL)`: Join multiple strings into a single string.

    ```{r}
    str_c(letters, LETTERS)
    ```

-   `str_flatten(string, collapse = "")`: Combines into a single string, separated by collapse.

    ```{r}
    str_flatten(fruit, ", ")
    ```

-   `str_dup(string, times)`: Repeat strings times times.
    Also `str_unique()` to remove duplicates.

    ```{r}
    str_dup(fruit, times = 2)
    ```

-   `str_split_fixed(string, pattern, n)`: Split a vector of strings into a matrix of substrings (splitting at occurrences of a pattern match).
    Also `str_split()` to return a list of substrings and `str_split_i()` to return the ith substring.

    ```{r}
    str_split_fixed(sentences, " ", n = 3)
    ```

-   `str_glue(..., .sep = "", .envir = parent.frame())`: Create a string from strings and {expressions} to evaluate.

    ```{r}
    str_glue("Pi is {pi}")
    ```

-   `str_glue_data(.x, ..., .sep = "", .envir = parent.frame(), .na = "NA")`: Use a data frame, list, or environment to create a string from strings and {expressions} to evaluate.

    ```{r}
    str_glue_data(mtcars, "{rownames(mtcars)} has {hp} hp")
    ```

## Manage Lengths

-   `str_length(string)`: The width of strings (i.e. number of code points, which generally equals the number of characters).

    ```{r}
    str_length(fruit)
    ```

-   `str_pad(string, width, side = c("left", "right", "both"), pad = " ")`: Pad strings to constant width.

    ```{r}
    str_pad(fruit, 17)
    ```

-   `str_trunc(string, width, side = c("left", "right", "both"), ellipsis = "...")`: Truncate the width of strings, replacing content with ellipsis.

    ```{r}
    str_trunc(sentences, 6)
    ```

-   `str_trim(string, side = c("left", "right", "both"))`: Trim whitespace from the start and/or end of a string.

    ```{r}
    str_trim(str_pad(fruit, 17))
    ```

-   `str_squish(string)`: Trim white space from each end and collapse multiple spaces into single spaces.

    ```{r}
    str_squish(str_pad(fruit, 17, "both"))
    ```

## Order Strings

-   `str_order(x, decreasing = FALSE, na_last = TRUE, locale = "en", numeric = FALSE, ...)^1^`: Return the vector of indexes that sorts a character vector.

    ```{r}
    fruit[str_order(fruit)]
    ```

-   `str_sort(x, decreasing = FALSE, na_last = TRUE, locale = "en", numeric = FALSE, ...)^1^`: Sort a character vector.

    ```{r}
    str_sort(fruit)
    ```

## Helpers

-   `str_conv(string, encoding)`: Override the encoding of a string.

    ```{r}
    str_conv(fruit, "ISO-8859-1")
    ```

-   `str_view(string, pattern, match = NA)`: View HTML rendering of all regex matches.
    Also `str_view()` to see only the first match.

    ```{r}
    str_view(sentences, "[aeiou]")
    ```

-   `str_equal(x, y, locale = "en", ignore_case = FALSE, ...)`^1^: Determine if two strings are equivalent.

    ```{r}
    str_equal(c("a", "b"), c("a", "c"))
    ```

-   `str_wrap(string, width = 80, indent = 0, exdent = 0)`: Wrap strings into nicely formatted paragraphs.

    ```{r}
    str_wrap(sentences, 20)
    ```

^1^ See <http://bit.ly/ISO639-1> for a complete list of locales.

<!-- Page 2 -->

## Regular Expressions

Regular expressions, or *regexps*, are a concise language for describing patterns in strings.

### Need to Know

Pattern arguments in stringr are interpreted as regular expressions *after any special characters have been parsed*.

In R, you write regular expressions as *strings*, sequences of characters surrounded by quotes(`""`) or single quotes (`''`).

Some characters cannot be directly represented in an R string.
These must be represented as **special characters**, sequences of characters that have a specific meaning, e.g. `\\` represents `\`, `\"` represents `"`, and `\n` represents a new line.
Run `?"'"` to see a complete list.

Because of this, whenever a `\` appears in a regular expression, you must write it as `\\` in the string that represents the regular expression.

Use `writeLines()` to see how R views your string after all special characters have been parsed.

For example, `writeLines("\\.")` will be parsed as `\.`

and `writeLines("\\ is a backslash")` will be parsed as `\ is a backslash`.

### Interpretation

Patterns in stringr are interpreted as regexs.
To change this default, wrap the pattern in one of:

-   `regex(pattern, ignore_case = FALSE, multiline = FALSE, comments = FALSE, dotall = FALSE, ...)`: Modifies a regex to ignore cases, match end of lines as well as end of strings, allow R comments within regexs, and/or to have `.` match everthing including `\n`.

    ```{r}
    str_detect("I", regex("i", TRUE))
    ```

-   `fixed()`: Matches raw bytes but will miss some characters that can be represented in multiple ways (fast).

    ```{r}
    str_detect("\u0130", fixed("i"))
    ```

-   `coll()`: Matches raw bytes and will use locale specific collation rules to recognize characters that can be represented in multiple ways (slow).

    ```{r}
    str_detect("\u0130", coll("i", TRUE, locale = "tr"))
    ```

-   `boundary()`: Matches boundaries between characters, line_breaks, sentences, or words.

    ```{r}
    str_split(sentences, boundary("word"))
    ```

### Match Characters

```{r}
see <- function(rx) str_view("abc ABC 123\t.!?\\(){}\n", rx)
```

+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
| string\     | regex\         | matches\                          | example              | example output (highlighted characters are in \<\>)               |
| (type this) | (to mean this) | (which matches this)              |                      |                                                                   |
+=============+================+===================================+======================+===================================================================+
|             | `a (etc.)`     | `a (etc.)`                        | `see("a")`           | ```                                                               |
|             |                |                                   |                      | <a>bc ABC 123\t.!?\(){}\n                                         |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
| `\\.`       | `\.`           | `.`                               | ``` see("\\.")`` ``` | ```                                                               |
|             |                |                                   |                      | abc ABC 123\t<.>!?\(){}\n                                         |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
| `\\!`       | `\!`           | `!`                               | `see("\\!")`         | ```                                                               |
|             |                |                                   |                      | abc ABC 123\t.<!>?\(){}\n                                         |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
| `\\?`       | `\?`           | `?`                               | `see("\\?")`         | ```                                                               |
|             |                |                                   |                      | abc ABC 123\t.!<?>\(){}\n                                         |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
| `\\\\`      | `\\`           | `\`                               | `see("\\\\")`        | ```                                                               |
|             |                |                                   |                      | abc ABC 123\t.!?<\>(){}\n                                         |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
| `\\(`       | `\(`           | `(`                               | `see("\\(")`         | ```                                                               |
|             |                |                                   |                      | abc ABC 123\t.!?\<(>){}\n                                         |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
| `\\)`       | `\)`           | `)`                               | `see("\\)")`         | ```                                                               |
|             |                |                                   |                      | abc ABC 123\t.!?\(<)>{}\n                                         |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
| `\\{`       | `\{`           | `{`                               | `see("\\{")`         | ```                                                               |
|             |                |                                   |                      | abc ABC 123\t.!?\()<{>}\n                                         |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
| `\\}`       | `\}`           | `}`                               | `see("\\}")`         | ```                                                               |
|             |                |                                   |                      | abc ABC 123\t.!?\(){<}>\n                                         |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
| `\\n`       | `\n`           | new line (return)                 | `see("\\n")`         | ```                                                               |
|             |                |                                   |                      | abc ABC 123\t.!?\(){}<\n>                                         |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
| `\\t`       | `\t`           | tab                               | `see("\\t")`         | ```                                                               |
|             |                |                                   |                      | abc ABC 123<\t>.!?\(){}\n                                         |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
| `\\s`       | `\s`           | any whitespace\                   | `see("\\s")`         | ```                                                               |
|             |                | (`\S` for non-whitespaces)        |                      | abc< >ABC< >123<\t>.!?\(){}<\n>                                   |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
| `\\d`       | `\d`           | any digit\                        | `see("\\d")`         | ```                                                               |
|             |                | (`\D` for non-digits)             |                      | abc ABC <1><2><3>\t.!?\(){}\n                                     |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
| `\\w`       | `\w`           | any word character\               | `see("\\w")`         | ```                                                               |
|             |                | (`\W` for non-word characters)    |                      | <a><b><c> <A><B><C> <1><2><3>\t.!?\(){}\n                         |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
| `\\b`       | `\b`           | word boundaries                   | `see("\\b")`         | ```                                                               |
|             |                |                                   |                      | <>abc<> <>ABC<> <>123<>\t.!?\(){}\n                               |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
|             | `[:digit:]`^1^ | digits                            | `see("[:digit:]")`   | ```                                                               |
|             |                |                                   |                      | abc ABC <1><2><3>\t.!?\(){}\n                                     |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
|             | `[:alpha:]`^1^ | letters                           | `see("[:alpha:]")`   | ```                                                               |
|             |                |                                   |                      | <a><b><c> <A><B><C> 123\t.!?\(){}\n                               |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
|             | `[:lower:]`^1^ | lowercase letters                 | `see("[:lower:]")`   | ```                                                               |
|             |                |                                   |                      | <a><b><c> ABC 123\t.!?\(){}\n                                     |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
|             | `[:upper:]`^1^ | uppercase letters                 | `see("[:upper:]")`   | ```                                                               |
|             |                |                                   |                      | abc <A><B><C> 123\t.!?\(){}\n                                     |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
|             | `[:alnum:]`^1^ | letters and numbers               | `see("[:alnum:]")`   | ```                                                               |
|             |                |                                   |                      | <a><b><c> <A><B><C> <1><2><3>\t.!?\(){}\n                         |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
|             | `[:punct:]`^1^ | punctuation                       | `see("[:punct:]")`   | ```                                                               |
|             |                |                                   |                      | abc ABC 123\t<.><!><?><\><(><)><{><}>\n                           |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
|             | `[:graph:]`^1^ | letters, numbers, and punctuation | `see("[:graph:]")`   | ```                                                               |
|             |                |                                   |                      | <a><b><c> <A><B><C> <1><2><3>\t<.><!><?><\><(><)><{><}>\n         |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
|             | `[:space:]`^1^ | space characters (i.e. `\s`)      | `see("[:space:]")`   | ```                                                               |
|             |                |                                   |                      | abc< >ABC< >123<\t>.!?\(){}<\n>                                   |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
|             | `[:blank:]`^1^ | space and tab (but not new line)  | `see("[:blank:]")`   | ```                                                               |
|             |                |                                   |                      | abc< >ABC< >123<\t>.!?\(){}\n                                     |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+
|             | `.`            | every character except a new line | `see(".")`           | ```                                                               |
|             |                |                                   |                      | <a><b><c>< ><A><B><C>< ><1><2><3><\t><.><!><?><\><(><)><{><}><\n> |
|             |                |                                   |                      | ```                                                               |
+-------------+----------------+-----------------------------------+----------------------+-------------------------------------------------------------------+

: 1Many base R functions require classes to be wrapped in a second set of \[ \], e.g. **\[\[:digit:\]\]**

#### Classes

-   The `[:space:]` class includes new line, and the `[:blank:]` class
    -   The `[:blank:]` class includes space and tab (`\t`)
-   The `[:graph:]` class contains all non-space characters, including `[:punct:]`, `[:symbol:]`, `[:alnum:]`, `[:digit:]`, `[:alpha:]`, `[:lower:]`, and `[:upper:]`
    -   `[:punct:]` contains punctuation: `. , : ; ? ! / * @ # - _ " [ ] { } ( )`

    -   `[:symbol:]` contains symbols: `` | ` = + ^ ~ < > $ ``

    -   `[:alnum:]` contains alphanumeric characters, including `[:digit:]`, `[:alpha:]`, `[:lower:]`, and `[:upper:]`

        -   `[:digit:]` contains the digits 0 through 9

        -   `[:alpha:]` contains letters, including `[:upper:]` and `[:lower:]`

            -   `[:upper:]` contains uppercase letters and `[:lower:]` contains lowercase letters
-   The regex `.` contains all characters in the above classes, except new line.

### Alternates

`alt <- function(rx) str_view("abcde", rx)`

+----------+--------------+-----------------+--------------------------------------+
| regexp   | matches      | example         | example output\                      |
|          |              |                 | (highlighted characters are in \<\>) |
+==========+==============+=================+======================================+
| `ab|d`   | or           | `alt("ab|d")`   | ```                                  |
|          |              |                 | <ab>c<d>e                            |
|          |              |                 | ```                                  |
+----------+--------------+-----------------+--------------------------------------+
| `[abe]`  | one of       | `alt("[abe]"`   | ```                                  |
|          |              |                 | <a><b>cd<e>                          |
|          |              |                 | ```                                  |
+----------+--------------+-----------------+--------------------------------------+
| `[^abe]` | anything but | `alt("[^abe]")` | ```                                  |
|          |              |                 | ab<c><d>e                            |
|          |              |                 | ```                                  |
+----------+--------------+-----------------+--------------------------------------+
| `[a-c]`  | range        | `alt("[a-c]")`  | ```                                  |
|          |              |                 | <a><b><c>de                          |
|          |              |                 | ```                                  |
+----------+--------------+-----------------+--------------------------------------+

: Alternates

### Anchors

`anchor <- function(rx) str_view("aaa", rx)`

+--------------------------------------------------------------------------------------------------------------------------------------------+
| regexp \| matches \| example \| example output\                                                                                            |
| \| \| \| (highlighted characters are in \<\>)                                                                                              |
+============================================================================================================================================+
| `^a` \| start of string \| `anchor("^a")` \| `|                 |                | <a>aa             |                 |                |` |
+--------------------------------------------------------------------------------------------------------------------------------------------+
| `a$` \| end of string \| `anchor("a$")` \| `|                 |                | aa<a>             |                 |                |`   |
+--------------------------------------------------------------------------------------------------------------------------------------------+

: Anchors

### Look Arounds

`look <- function(rx) str_view("bacad", rx)`

+-----------+-----------------+-------------------+--------------------------------------+
| regexp    | matches         | example           | example output\                      |
|           |                 |                   | (highlighted characters are in \<\>) |
+===========+=================+===================+======================================+
| `a(?=c)`  | followed by     | `look("a(?=c)")`  | ```                                  |
|           |                 |                   | b<a>cad                              |
|           |                 |                   | ```                                  |
+-----------+-----------------+-------------------+--------------------------------------+
| `a(?!c)`  | not followed by | `look("a(?!c)")`  | ```                                  |
|           |                 |                   | bac<a>d                              |
|           |                 |                   | ```                                  |
+-----------+-----------------+-------------------+--------------------------------------+
| `(?<=b)a` | preceded by     | `look("(?<=b)a")` | ```                                  |
|           |                 |                   | b<a>cad                              |
|           |                 |                   | ```                                  |
+-----------+-----------------+-------------------+--------------------------------------+
| `(?<!b)a` | not preceded by | `look("(?<!b)a")` | ```                                  |
|           |                 |                   | bac<a>d                              |
|           |                 |                   | ```                                  |
+-----------+-----------------+-------------------+--------------------------------------+

: Look arounds

### Quantifiers

`quant <- function(rx) str_view(".a.aa.aaa", rx)`

+-----------+---------------------+-------------------+--------------------------------------+
| regexp    | matches             | example           | example output\                      |
|           |                     |                   | (highlighted characters are in \<\>) |
+===========+=====================+===================+======================================+
| `a?`      | zero or one         | `quant("a?")`     | ```                                  |
|           |                     |                   | <>.<a><>.<a><a><>.<a><a><a><>        |
|           |                     |                   | ```                                  |
+-----------+---------------------+-------------------+--------------------------------------+
| `a*`      | zero or more        | `quant("a*")`     | ```                                  |
|           |                     |                   | <>.<a><>.<aa><>.<aaa><>              |
|           |                     |                   | ```                                  |
+-----------+---------------------+-------------------+--------------------------------------+
| `a+`      | one or more         | `quant("a+")`     | ```                                  |
|           |                     |                   | .<a>.<aa>.<aaa>                      |
|           |                     |                   | ```                                  |
+-----------+---------------------+-------------------+--------------------------------------+
| `a{n}`    | exactly `n`         | `quant("a{2}")`   | ```                                  |
|           |                     |                   | .a.<aa>.<aa>a                        |
|           |                     |                   | ```                                  |
+-----------+---------------------+-------------------+--------------------------------------+
| `a{n, }`  | `n` or more         | `quant("a{2,}")`  | ```                                  |
|           |                     |                   | .a.<aa>.<aaa>                        |
|           |                     |                   | ```                                  |
+-----------+---------------------+-------------------+--------------------------------------+
| `a{n, m}` | between `n` and `m` | `quant("a{2,4}")` | ```                                  |
|           |                     |                   | .a.<aa>.<aaa>                        |
|           |                     |                   | ```                                  |
+-----------+---------------------+-------------------+--------------------------------------+

: Quantifiers

### Groups

`ref <- function(rx) str_view("abbaab", rx)`

Use parentheses to set precedent (order of evaluation) and create groups

+-----------+-----------------+------------------+--------------------------------------+
| regexp    | matches         | example          | example output\                      |
|           |                 |                  | (highlighted characters are in \<\>) |
+===========+=================+==================+======================================+
| `(ab|d)e` | sets precedence | `alt("(ab|d)e")` | ```                                  |
|           |                 |                  | abc<de>                              |
|           |                 |                  | ```                                  |
+-----------+-----------------+------------------+--------------------------------------+

: Groups

Use an escaped number to refer to and duplicate parentheses groups that occur earlier in a pattern.
Refer to each group by its order of appearance

+-------------+----------------+----------------------+-------------------------------------------+--------------------------------------+
| string\     | regexp\        | matches\             | example\                                  | example output\                      |
| (type this) | (to mean this) | (which matches this) | (the result is the same as `ref("abba")`) | (highlighted characters are in \<\>) |
+=============+================+======================+===========================================+======================================+
| `\\1`       | `\1` (etc.)    | first () group, etc. | `ref("(a)(b)\\2\\1")`                     | ```                                  |
|             |                |                      |                                           | <abba>ab                             |
|             |                |                      |                                           | ```                                  |
+-------------+----------------+----------------------+-------------------------------------------+--------------------------------------+

: More groups

------------------------------------------------------------------------

CC BY SA Posit Software, PBC • [info\@posit.co](mailto:info@posit.co) • [posit.co](https://posit.co)

Learn more at [stringr.tidyverse.org](https://stringr.tidyverse.org).

Updated: `r format(Sys.Date(), "%Y-%m")`.

```{r}
#| output: true

packageVersion("stringr")
```

------------------------------------------------------------------------
