---
title: "Data tidying with tidyr :: Cheatsheet"
description: " "
image-alt: ""
execute:
  eval: true
  output: false
  warning: false
editor_options: 
  chunk_output_type: console
---

```{r}
#| output: asis
#| echo: false
#| column: margin
source("common.R")
use_cheatsheet_logo(
  "tidyr", 
  alt = "Hex logo for tidyr - three different brightly coloured cartoon brooms that appear as spaceships flying through a black background, with coloured spots representing stars and planets.")
sheet_name <- tools::file_path_sans_ext(knitr::current_input())
pdf_preview_link(sheet_name)
translation_list(sheet_name)
```

<!-- Page 1 -->

**Tidy data** is a way to organize tabular data in a consistent data structure across packages.
A table is tidy if:

-   Each **variable** is in its own **column**
-   Each **observation**, or **case**, is in its own **row**
-   Access **variables** as **vectors**
-   Preserve **cases** in vectorized operations

```{r}
library(tidyr)
library(tibble)
```

```{r}
#| include: false

library(dplyr)
```

## Tibbles

### An Enhanced Data Frame

Tibbles are a table format provided by the **tibble** package.
They inherit the data frame class, but have improved behaviors:

-   **Subset** a new tibble with `]`, a vector with `[[` and `$`.

-   **No partial matching** when subsetting columns.

-   **Display** concise views of the data on one screen.

-   `options(tibble.print_max = n, tibble.print_min = m, tibble.width = Inf)`: Control default display settings.

-   `View()` or `glimpse()`: View the entire data set.

### Construct a Tibble

-   `tibble(...)`: Construct by columns.

    ```{r}
    #| output: true

    tibble(
      x = 1:3, 
      y = c("a", "b", "c")
    )
    ```

-   `tribble(...)`: Construct by rows.

    ```{r}
    tribble(
      ~x, ~y,
       1, "a",
       2, "b",
       3, "c"
    )
    ```

-   `as_tibble(x, ...)`: Convert a data frame to a tibble.

-   `enframe(x, name = "name", value = "value")`: Convert a named vector to a tibble.
    Also `deframe()`.

-   `is_tibble(x)`: Test whether x is a tibble.

## Reshape Data

Pivot data to reorganize values into a new layout.

-   `pivot_longer(data, cols, name_to = "name", values_to = "value", values_drop_na = FALSE)`: "Lengthen" data by collapsing several columns into two.

    -   The initial `table4a` looks like the following:

    ```{r}
    #| output: true

    table4a
    ```

    -   Column names move to a new `names_to` column and values to a new `values_to` column. The output of `pivot_longer()` will look like the following:

    ```{r}
    #| output: true

    pivot_longer(table4a, cols = 2:3, names_to = "year", values_to = "cases")
    ```

-   `pivot_wider(data, name_from = "name", values_from = "value")`: The inverse of `pivot_longer()`.
    "Widen" data by expanding two columns into several.

    -   The initial `table2` looks like the following:

    ```{r}
    #| output: true

    table2
    ```

    -   One column provides the new column names, the other the values. The output of `pivot_wider()` will look like the following:

    ```{r}
    #| output: true

    pivot_wider(table2, names_from = type, values_from = count)
    ```

## Split Cells

Use these functions to split or combine cells into individual, isolated values.

-   `unite(data, col, ..., sep = "_", remove = TRUE, na.rm = FALSE)`: Collapse cells across several columns into a single column.

    -   The initial `table5` looks like the following:

    ```{r}
    #| output: true

    table5
    ```

    -   The output of `unite()` will look like the following:

    ```{r}
    #| output: true

    unite(table5, century, year, col = "year", sep = "")
    ```

-   `separate_wider_delim(data, cols, delim, ..., names = NULL, names_sep = NULL, names_repair = "check unique", too_few, too_many, cols_remove = TRUE)`: Separate each cell in a column into several columns.
    Also `extract()`.

    -   The initial `table3` looks like the following:

    ```{r}
    #| output: true

    table3
    ```

    -   The output of `separate_wider_delim()` will look like the following:

    ```{r}
    #| output: true

    separate_wider_delim(table3, rate, delim = "/", names = c("cases", "pop"))
    ```

-   `separate_longer_delim(data, cols, delim, .., width, keep_empty)`: Separate each cell in a column into several rows.

    -   The initial `table3` looks like the following:

    ```{r}
    #| output: true

    table3
    ```

    -   The output of `separate_longer_delim()` will look like the following:

    ```{r}
    #| output: true

    separate_longer_delim(table3, rate, delim = "/")
    ```

## Expand Tables

Create new combinations of variables or identify implicit missing values (combinations of variables not present in the data).

-   `expand(data, ...)`: Create a new tibble with all possible combinations of the values of the variables listed in ...
    Drop other variables.

    ```{r}
    expand(mtcars, cyl, gear, carb)
    ```

-   `complete(data, ..., fill = list())`: Add missing possible combinations of values of variables listed in ...
    Fill remaining variables with NA.

    ```{r}
    complete(mtcars, cyl, gear, carb)
    ```

## Handle Missing Values

Drop or replace explicit missing values (`NA`).

```{r}
x <- tribble(
  ~x1, ~x2,
  "A",   1,
  "B",  NA,
  "C",  NA,
  "D",   3,
  "E",  NA
)
```

-   `drop_na(data, ...)`: Drop rows containing `NA`s in ...
    columns.

    ```{r}
    drop_na(x, x2)
    ```

-   `fill(data, ..., .direction = "down")`: Fill in `NA`s in ...
    columns using the next or previous value.

    ```{r}
    fill(x, x2)
    ```

-   `replace_na(data, replace)`: Specify a value to replace `NA` in selected columns.

    ```{r}
    replace_na(x, list(x2 = 2))
    ```

<!-- Page 2 -->

## Nested Data

A **nested data frame** stores individual tables as a list-column of data frames within a larger organizing data frame.
List-columns can also be lists of vectors or lists of varying data types.
Use a nested data frame to:

-   Preserve relationships between observations and subsets of data. Preserve the type of the variables being nested (factors and datetimes aren't coerced to character).
-   Manipulate many sub-tables are once with **purrr** functions like `map()`, `map2()`, or `pmap()` or with **dplyr** `rowwise()` grouping.

### Create Nested Data

-   `nest(data, ...)`: Moves groups of cells into a list-column of a data frame. Use alone or with `dplyr::group_by()`.

1.  Group the data frame with `group_by()` and use `nest()` to move the groups into a list-column.

    ```{r}
    n_storms <- storms |>
      group_by(name) |>
      nest()
    ```

2.  Use `nest(new_col = c(x,y))` to specify the columns to group using `dplyr::select()` syntax.

    ```{r}
    n_storms <- storms |>
      nest(data = c(year:long))
    ```

-   Index list-columns with `[[]]`.

    ```{r}
    n_storms$data[[1]]
    ```

### Create Tibbles With List-Columns

-   `tibble::tribble(...)`: Makes list-columns when needed.

    ```{r}
    tribble(
      ~max, ~seq,
         3,  1:3,
         4,  1:4,
         5,  1:5
    )
    ```

-   `tibble::tibble(...)`: Saves list input as list-columns.

    ```{r}
    tibble(
      max = c(3,4,5),
      seq = list(1:3, 1:4, 1:5)
    )
    ```

-   `tibble::enframe(x, name = "name", value = "value")`: Convert multi-level list to a tibble with list-cols.

    ```{r}
    enframe(list("3" = 1:3, "4" = 1:4, "5" = 1:5), "max", "seq")
    ```

### Output List-Columns From Other Functions

-   `dplyr::mutate()`, `transmute()`, and `summarise()` will output list-columns if they return a list.

    ```{r}
    mtcars |>
      group_by(cyl) |>
      summarise(q = list(quantile(mpg)))
    ```

### Reshape Nested Data

-   `unnest(data, cols, ..., keep_empty = FALSE)`: Flatten nested columns back to regular columns.
    The inverse of `nest()`.

    ```{r}
    n_storms |> unnest(data)
    ```

-   `unnest_longer(data, col, values_to = NULL, indices_to = NULL)`: Turn each element of a list-column into a row.

    ```{r}
    starwars |>
      select(name, films) |>
      unnest_longer(films)
    ```

-   `unnest_wider(data, col)`: Turn each element of a list-column into a regular column.

    ```{r}
    starwars |>
      select(name, films) |>
      unnest_wider(films, names_sep = "_")
    ```

-   `hoist(.data, .col, ..., remove = TRUE)`: Selectively pull list components out into their own top-level columns.
    Uses `purrr::pluck()` syntax for selecting from lists.

    ```{r}
    starwars |>
      select(name, films) |>
      hoist(films, first_film = 1, second_film = 2)
    ```

### Transform Nested Data

A vectorized function takes a vector, transforms each element in parallel, and returns a vector of the same length.
By themselves vectorized functions cannot work with lists, such as list-columns.

-   `dplyr::rowwise(.data, ...)`: Group data so that each row is one group, and within the groups, elements of list-columns appear directly (accessed with `[[`), not as lists of length one.
    **When you use rowwise(), dplyr functions will seem to apply functions to list-columns in a vectorized fashion.**

-   Apply a function to a list-column and **create a new list-column.** In this example, `dim()` returns two values per row and so is wrapped with `list()` to tell `mutate()` to create a list-column.

    ```{r}
    n_storms |>
      rowwise() |>
      mutate(n = list(dim(data))) # dim() returns two values per row, wrap with list to tell mutate to create a list-column
    ```

-   Apply a function to a list-column and **create a regular column.** In this example, `nrow()` returns one integer per row.

    ```{r}
    n_storms |>
      rowwise() |>
      mutate(n = nrow(data)) # nrow() returns one integer per row
    ```

-   Collapse **multiple list-columns** into a single list-column.
    In this example, `append()` returns a list for each row, so col type must be list.

    ```{r}
    starwars |>
      rowwise() |>
      mutate(transport = list(append(vehicles, starships))) # append() returns a list for each row, so col type must be list
    ```

-   Apply a function to **multiple list-columns.** In this example, `length()` returns one integer per row.

    ```{r}
    starwars |>
      rowwise() |>
      mutate(n_transports = length(c(vehicles, starships)))
      # length() returns one integer per row
    ```

-   See **purrr** package for more list functions.

------------------------------------------------------------------------

CC BY SA Posit Software, PBC • [info\@posit.co](mailto:info@posit.co) • [posit.co](https://posit.co)

Learn more at [tidyr.tidyverse.org](https://tidyr.tidyverse.org).

Updated: `r format(Sys.Date(), "%Y-%m")`.

```{r}
#| output: true

packageVersion("tidyr")
packageVersion("tibble")
```

------------------------------------------------------------------------
