---
title: "Apply functions with purrr :: Cheat Sheet"
format: 
  html:
    toc: true
    highlight-style: a11y-dark
    css: css/cheatsheet.css
editor: visual
execute:
  eval: true
  output: false
  message: true
---

purrr enhances R's functional programming (FP) toolkit by providing a complete and consistent set of tools for working with functions and vectors.
If you've never heard of FP before, the best place to start is the family of [`map()`](https://purrr.tidyverse.org/reference/map.html) functions which allow you to replace many for loops with code that is both more succinct and easier to read.
The best place to learn about the [`map()`](https://purrr.tidyverse.org/reference/map.html) functions is the [iteration chapter](https://r4ds.had.co.nz/iteration.html) in R for Data Science.

```{r}
library(purrr)
```

<!-- Page 1 -->

## Map Functions

### One List

-   `map(.x, .f, ...)`: Apply a function to each element of a list of vector, and return a list.

    ```{r}
    x <- list(a = 1:10, b = 11:20, c = 21:30)
    l1 <- list(x = c("a", "b"), y = c("c", "d"))
    map(l1, sort, decreasing = TRUE)
    ```

-   `map_dbl(.x, .f, ...)`: Return a double vector.

    ```{r}
    map_dbl(x, mean)
    ```

-   `map_int(.x, .f, ...)`: Return an integral vector.

    ```{r}
    map_int(x, length)
    ```

-   `map_chr(.x, .f, ...)`: Return a character vector.

    ```{r}
    map_chr(l1, paste, collapse = "")
    ```

-   `map_lgl(.x, .f, ...)`: Return a logical vector.

    ```{r}
    map_lgl(x, is.integer)
    ```

-   `map_dfc(.x, .f, ...)`: Return a data frame created by column-binding.

    ```{r}
    map_dfc(l1, rep, 3)
    ```

-   `map_dfr(.x, .f, ...)`: Return a data frame created by row-binding.

    ```{r}
    map_dfr(x, summary)
    ```

-   `walk(.x, .f, ...)`: Trigger side effects, return invisibly.

    ```{r}
    walk(x, print)
    ```

### Two Lists

-   `map2(.x, .y, .f, ...)`: Apply a function pairs of elements from two lists or vectors, return a list.

    ```{r}
    y <- list(1, 2, 3)
    z <- list(4, 5, 6)
    l2 <- list(x = "a", y = "z")
    map2(x, y, ~ .x * .y)
    ```

-   `map2_dbl(.x, .y, .f, ...)`: Return a double vector.

    ```{r}
    map2_dbl(y, z, ~ .x / .y)
    ```

-   `map2_int(.x, .y, .f, ...)`: Return an integral vector.

    ```{r}
    map2_int(y, z, `+`)
    ```

-   `map2_chr(.x, .y, .f, ...)`: Return a character vector.

    ```{r}
    map2_chr(l1, l2, paste, collapse = ",", sep = ":")
    ```

-   `map2_lgl(.x, .y, .f, ...)`: Return a logical vector.

    ```{r}
    map2_lgl(l2, l1, `%in%`)
    ```

-   `map2_dfc(.x, .y, .f, ...)`: Return a data frame created by column-binding.

    ```{r}
    map2_dfc(l1, l2, ~ as.data.frame(c(.x, .y)))
    ```

-   `map2_dfr(.x, .y, .f, ...)`: Return a data frame created by row-binding.

    ```{r}
    map2_dfr(l1, l2, ~ as.data.frame(c(.x, .y)))
    ```

-   `walk2(.x, .y, .f, ...)`: Trigger side effects, return invisibly.

    ```{r}
    #| eval: false

    walk2(objs, paths, save)
    ```

### Many Lists

-   `pmap(.l, .f, ...)`: Apply a function to groups of elements from a list of lists or vectors, return a list.

    ```{r}
    pmap(list(x, y, z), ~ ..1 * (..2 + ..3))
    ```

-   `pmap_dbl(.l, .f, ...)`: Return a double vector.

    ```{r}
    pmap_dbl(list(y, z), ~ .x / .y)
    ```

-   `pmap_int(.l, .f, ...)`: Return an integral vector.

    ```{r}
    pmap_int(list(y, z), `+`)
    ```

-   `pmap_chr(.l, .f, ...)`: Return a character vector.

    ```{r}
    pmap_chr(list(l1, l2), paste, collapse = ",", sep = ":")
    ```

-   `pmap_lgl(.l, .f, ...)`: Return a logical vector.

    ```{r}
    pmap_lgl(list(l2, l1), `%in%`)
    ```

-   `pmap_dfc(.l, .f, ...)`: Return a data frame created by column-binding.

    ```{r}
    pmap_dfc(list(l1, l2), ~ as.data.frame(c(.x, .y)))
    ```

-   `pmap_dfr(.l, .f, ...)`: Return a data frame created by row-binding.

    ```{r}
    pmap_dfr(list(l1, l2), ~ as.data.frame(c(.x, .y)))
    ```

-   `pwalk(.l, .f, ...)`: Trigger side effects, return invisibly.

    ```{r}
    #| eval: false

    pwalk(list(objs, paths), save)
    ```

### Lists and Indexes

-   `imap(.x, .f, ...)`: Apply `.f` to each element and its index, return a list.

    ```{r}
    imap(y, ~ paste0(.y, ": ", .x))
    ```

-   `imap_dbl(.x, .f, ...)`: Return a double vector.

    ```{r}
    imap_dbl(y, ~ .y)
    ```

-   `imap_int(.x, .f, ...)`: Return an integral vector.

    ```{r}
    imap_int(y, ~ .y)
    ```

-   `imap_chr(.x, .f, ...)`: Return a character vector.

    ```{r}
    imap_chr(y, ~ paste0(.y, ": ", .x))
    ```

-   `imap_lgl(.x, .f, ...)`: Return a logical vector.

    ```{r}
    imap_lgl(l1, ~ is.character(.y))
    ```

-   `imap_dfc(.x, .f, ...)`: Return a data frame created by column-binding.

    ```{r}
    imap_dfc(l2, ~ as.data.frame(c(.x, .y)))
    ```

-   `imap_dfr(.x, .f, ...)`: Return a data frame created by row-binding.

    ```{r}
    imap_dfr(l2, ~ as.data.frame(c(.x, .y)))
    ```

-   `iwalk(.x, .f, ...)`: Trigger side effects, return invisibly.

    ```{r}
    iwalk(z, ~ print(paste0(.y, ": ", .x)))
    ```

## Function Shortcuts

-   Use `~ .` with functions like `map()` that have single arguments.
    `map(l, ~ . + 2)` becomes `map(l, function(x) x + 2)`.

-   Use `~ .x .y` with functions like `map2()` that have two arguments.
    `map2(l, p, ~ .x + .y)` becomes `map2(l, p, function(l, p) l + p)`.

-   Use `~ ..1 ..2 ..3` etc with functions like `pmap()` that have many arguments.
    `pmap(list(a, b, c), ~ ..3 + ..1 - ..2)` becomes `pmap(list(a,b,c), function(a, b, c) c + a - b)`.

-   Use `~ .x .y` with functions like `imap()`.
    `.x` will get the list value and `.y` with get the index, or name if available.
    `imap(list(a, b, c), ~ paste0(.y, ": ", .x))` outputs `index: value` for each item.

-   Use a `string` or `integer` with any map function to index list elements by name or position.
    `map(l, "name")` becomes `map(l, function(x) x[["name"]])`.

<!-- Page 2 -->

## Work with Lists

### Filter

-   `keep(.x, .p, ...)`: Select elements that pass a logical test.
    Conversely `discard()`.

    ```{r}
    keep(x, is.numeric)
    ```

-   `compact(.x, .p = identity)`: Drop empty elements.

    ```{r}
    compact(x)
    ```

-   `head_while(.x, .p, ...)`: Return head elements until one does not pass.
    Also `tail_while()`.

    ```{r}
    head_while(x, is.character)
    ```

-   `detect(.x, .f, ..., dir = c("forward", "backward"), .right = NULL)`: Find first element to pass.

    ```{r}
    detect(x, is.character)
    ```

-   `detect_index(.x, .f, ..., dir = c("forward", "backward"), .right = NULL)`: Find index of first element to pass.

    ```{r}
    detect_index(x, is.character)
    ```

-   `every(.x, .p, ...)`: Do all elements pass a test?

    ```{r}
    every(x, is.character)
    ```

-   `some(.x, .p, ...)`: Do some elements pass a test?

    ```{r}
    some(x, is.character)
    ```

-   `none(.x, .p, ...)`: Do no elements pass a test?

    ```{r}
    none(x, is.character)
    ```

-   `has_element(.x, .y)`: Does a list contain an element?

    ```{r}
    has_element(x, "foo")
    ```

-   `pluck_depth(x)`: Return depth (number of levels of indexes).

    ```{r}
    pluck_depth(x)
    ```

### Index

-   `pluck(.x, ..., .deault = NULL)`: Select an element by name or index.
    Also `attr_getter()` and `chuck()`.

    ```{r}
    pluck(x, "b")
    x |> pluck("b")
    ```

-   `assign_in(x, where, value)`: Assign a value to a location using pluck selection.

    ```{r}
    assign_in(x, "b", 5)
    x |> assign_in("b", 5)
    ```

-   `modify_in(.x, .where,, .f)`: Apply a function to a value at a selected location.

    ```{r}
    modify_in(x, "b", abs)
    ```

### Reshape

-   `flatten(.x)`: Remove a level of indexes from a list.
    Also `flatten_chr()` etc.

    ```{r}
    flatten(x)
    ```

-   `array_tree(array, margin = NULL)`: Turn array into list.
    Also `array_branch()`.

    ```{r}
    z <- array(1:12, c(2, 2, 2))
    array_tree(z, margin = 3)
    ```

-   `transpose(.l, .names = NULL)`: Transposes the index order in a multi-level list.

    ```{r}
    transpose(x)
    ```

-   `set_names(x, nm = x)`: Set the names of a vector/list directly or with a function.

    ```{r}
    set_names(x, c("p", "q", "r"))
    set_names(x, tolower)
    ```

### Modify

-   `modify(.x, .f, ...)`: Apply a function to each element.
    Also `modify2()` and `imodify()`.

    ```{r}
    modify(x, ~ . + 2)
    ```

-   `modify_at(.x, .at, .f, ...)`: Apply a function to selected elements.
    Also `map_at()`.

    ```{r}
    modify_at(x, "b", ~ . + 2)
    ```

-   `modify_if(.x, .p, .f, ...)`: Apply a function to elements that pass a test.
    Also `map_if()`.

    ```{r}
    modify_if(x, is.numeric, ~ . + 2)
    ```

-   `modify_depth(.x, .depth, .f, ...)`: Apply function to each element at a given level of a list.
    Also `map_depth()`.

    ```{r}
    modify_depth(x, 1, ~ . + 2)
    ```

### Reduce

-   `reduce(.x, .f, ..., .init, .dir = c("forward", "backward"))`: Apply function recursively to each element of a list of vector.
    Also `reduce2()`.

    ```{r}
    a <- list(1, 2, 3, 4)
    reduce(a, sum)
    ```

-   `accumulate(.x, .f, ..., .init)`: Reduce a list, but also return intermediate results in a list.
    Also `accumulate2()`.

    ```{r}
    a <- list(1, 2, 3, 4)
    accumulate(a, sum)
    ```

### List-Columns

**List-columns** are columns of a data frame where each element is a list or vector instead of an atomic value.
Columns can also be lists of data frames.
See **tidyr** for more about nested data and list columns.

#### Work With List-Columns

Manipulate list-columns like any other kind of column, using **dplyr** functions like `mutate()` and `transmute()`.
Because each element is a list, use **map functions** within a column function to manipulate each element.

-   `map()`, `map2()`, or `pmap()` return lists and will **create new list-columns**.
    In this example, `transmute()` is a column function, `map2()` is a list function which returns a list, and `vehicles` and `starships` are list-columns.

    ```{r}
    dplyr::starwars |>
      dplyr::mutate(ships = map2(vehicles, starships, append))
    ```

-   Suffixed map functions like `map_int()` return an atomic data type and will **simplify list-columns into regular columns**.
    In this example, `mutate()` is a column function, `map_int()` is a list function which returns a column vector, and `films` is a list column.

    ```{r}
    dplyr::starwars |>
      dplyr::mutate(n_films = map_int(films, length))
    ```

------------------------------------------------------------------------

CC BY SA Posit Software, PBC • [info\@posit.co](mailto:info@posit.co) • [posit.co](https://posit.co)

Learn more at [purrr.tidyverse.org](https://purrr.tidyverse.org).

Updated: `r format(Sys.Date(), "%Y-%m")`.

```{r}
#| output: true

packageVersion("purrr")
```

------------------------------------------------------------------------

